---
title: "gnomAD vs NYGC vs TOPMed"
author: "Andy Beck"
date: "2023-03-07"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

In this document I will compare the results from running our pipeline on datasets from

1. The NYGC
2. gnomAD
3. Internal (TOPMed)

These three datasets all contain versions of the 1KGP deep-sequence release, with the difference being what panel/reference(?) they were called against. In all three cases, the VCFs were filtered to remove:

1. Samples other than the 2,406 unrelated subjects present in the gnomAD release
2. Variants other than snps
3. SNPs in either pseudo-autosomal region
4. Variants outside the 1KGP pilot accessibility mask
5. Remaining singleton variants.

```{r}
library(tidyverse)
library(janitor)
source('code/common_functions.R')

pair_info_df <- read_delim("data/sample_pairs_gnomad.csv", col_names = c("POP", "ID1", "ID2"))
pair_info_df$SP <- c(rep("EUR", 200), rep("AFR", 200))

gc_content_1kb <- read_tsv("data/gc1kb_X_only.bed")
colnames(gc_content_1kb) <- c("CHR", "START", "END", "AT", "GC", "A", "C", "G", "T", "TOTAL", "OTHER", "LENGTH")
gc_content_1kb  <- gc_content_1kb %>%
  mutate(bin_id = (START / 1000) + 1)
```

Loading the switch location files from vcftools:

```{r}
# Directories: TOPMed
base_dir_int <- "/net/snowwhite/home/beckandy/research/phasing/output/filter_switch_errors/switch_errors/"
eagle_dir_int <- paste0(base_dir_int, "eagle/annotated/")
bagle_dir_int <- paste0(base_dir_int, "beagle/annotated/")
shapeit_dir_int <- paste0(base_dir_int, "shapeit/annotated/")

# Directories: NYGC
base_dir_nygc <- "/net/snowwhite/home/beckandy/research/phasing/output/nygc/switch_errors/"
eagle_dir_nygc <- paste0(base_dir_nygc, "eagle/annotated/")
bagle_dir_nygc <- paste0(base_dir_nygc, "beagle/annotated/")
shapeit_dir_nygc <- paste0(base_dir_nygc, "shapeit/annotated/")

# Directories: gnomAD
base_dir_gnomad <- "/net/snowwhite/home/beckandy/research/phasing/output/gnomad/switch_errors/"
eagle_dir_gnomad <- paste0(base_dir_gnomad, "eagle/annotated/")
bagle_dir_gnomad <- paste0(base_dir_gnomad, "beagle/annotated/")
shapeit_dir_gnomad <- paste0(base_dir_gnomad, "shapeit/annotated/")

# Data frames
df_vcftools_int <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, eagle_dir_int, bagle_dir_int, shapeit_dir_int, gc_content_1kb)
                      }) %>%
  bind_rows()
df_vcftools_int$pop <- c(rep("EUR", 200), rep("AFR", 200))

## NYGC
df_vcftools_nygc <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, eagle_dir_nygc, bagle_dir_nygc, shapeit_dir_nygc, gc_content_1kb)
                      }) %>%
  bind_rows()
df_vcftools_nygc$pop <- c(rep("EUR", 200), rep("AFR", 200))

## gnomAD
df_vcftools_gnomad <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, eagle_dir_gnomad, bagle_dir_gnomad, shapeit_dir_gnomad, gc_content_1kb)
                      }) %>%
  bind_rows()
df_vcftools_gnomad$pop <- c(rep("EUR", 200), rep("AFR", 200))
```

And now from whatshap:

```{r}
whatshap_dir_int <- "/net/snowwhite/home/beckandy/research/phasing/output/filter_switch_errors/whatshap"
df_wh_eagle_int <- get_all_whatshap("eagle", whatshap_dir_int)
df_wh_beagle_int <- get_all_whatshap("beagle", whatshap_dir_int)
df_wh_si_int <- get_all_whatshap("shapeit", whatshap_dir_int)

whatshap_dir_nygc <- "/net/snowwhite/home/beckandy/research/phasing/output/nygc/whatshap"
df_wh_eagle_nygc <- get_all_whatshap("eagle", whatshap_dir_nygc)
df_wh_beagle_nygc <- get_all_whatshap("beagle", whatshap_dir_nygc)
df_wh_si_nygc <- get_all_whatshap("shapeit", whatshap_dir_nygc)

whatshap_dir_gnomad <- "/net/snowwhite/home/beckandy/research/phasing/output/gnomad/whatshap"
df_wh_eagle_gnomad <- get_all_whatshap("eagle", whatshap_dir_gnomad)
df_wh_beagle_gnomad <- get_all_whatshap("beagle", whatshap_dir_gnomad)
df_wh_si_gnomad <- get_all_whatshap("shapeit", whatshap_dir_gnomad)
```

## Flips and Swtiches Comparison

### SHAPEIT

#### gnomAD and TOPMed

```{r}
df_wh_si_gnomad %>%
  select(id, pop, flips) %>%
  rename(gnomAD = flips) %>%
  inner_join({df_wh_si_int %>%
      select(id, pop, flips) %>%
      rename(TOPMed = flips)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flips: Internal") +
  ylab("Flips: gnomAD") +
  ggtitle("Flips: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)

df_wh_si_gnomad %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_si_int %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(TOPMed = p_flip)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flip rate: Internal") +
  ylab("Flip rate: gnomAD") +
  ggtitle("Flips / Number of Heterozygous Sites: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)
```

```{r}
df_wh_si_gnomad %>%
  select(id, pop, switches) %>%
  rename(gnomAD = switches) %>%
  inner_join({df_wh_si_int %>%
      select(id, pop, switches) %>%
      rename(TOPMed = switches)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switches: Internal") +
  ylab("Switches: gnomAD") +
  ggtitle("Switches: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)

df_wh_si_gnomad %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_si_int %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(TOPMed = p_flip)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switch rate: Internal") +
  ylab("Switch rate: gnomAD") +
  ggtitle("Switches / Number of Heterozygous Sites: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)
```


#### gnomAD and NYGC

```{r}
df_wh_si_gnomad %>%
  select(id, pop, flips) %>%
  rename(gnomAD = flips) %>%
  inner_join({df_wh_si_nygc %>%
      select(id, pop, flips) %>%
      rename(NYGC = flips)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flips: NYGC") +
  ylab("Flips: gnomAD") +
  ggtitle("Flips: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)

df_wh_si_gnomad %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_si_nygc %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flip rate: NYGC") +
  ylab("Flip rate: gnomAD") +
  ggtitle("Flips / Number of Heterozygous Sites: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_si_gnomad %>%
  select(id, pop, switches) %>%
  rename(gnomAD = switches) %>%
  inner_join({df_wh_si_nygc %>%
      select(id, pop, switches) %>%
      rename(NYGC = switches)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switches: NYGC") +
  ylab("Switches: gnomAD") +
  ggtitle("Switches: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)

df_wh_si_gnomad %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_si_nygc %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switch rate: NYGC") +
  ylab("Switch rate: gnomAD") +
  ggtitle("Switches / Number of Heterozygous Sites: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)  
```

#### TOPMed and NYGC

```{r}
df_wh_si_int %>%
  select(id, pop, flips) %>%
  rename(TOPMed = flips) %>%
  inner_join({df_wh_si_nygc %>%
      select(id, pop, flips) %>%
      rename(NYGC = flips)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Flips: NYGC") +
  ylab("Flips: TOPMed") +
  ggtitle("Flips: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)

df_wh_si_int %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(TOPMed = p_flip) %>%
  inner_join({df_wh_si_nygc %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Flip rate: NYGC") +
  ylab("Flip rate: TOPMed") +
  ggtitle("Flips / Number of Heterozygous Sites: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_si_int %>%
  select(id, pop, switches) %>%
  rename(TOPMed = switches) %>%
  inner_join({df_wh_si_nygc %>%
      select(id, pop, switches) %>%
      rename(NYGC = switches)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Switches: NYGC") +
  ylab("Switches: TOPMed") +
  ggtitle("Switches: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)

df_wh_si_int %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(TOPMed = p_flip) %>%
  inner_join({df_wh_si_nygc %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Switch rate: NYGC") +
  ylab("Switch rate: TOPMed") +
  ggtitle("Switches / Number of Heterozygous Sites: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)  
```

### EAGLE

#### gnomAD and TOPMed

```{r}
df_wh_eagle_gnomad %>%
  select(id, pop, flips) %>%
  rename(gnomAD = flips) %>%
  inner_join({df_wh_eagle_int %>%
      select(id, pop, flips) %>%
      rename(TOPMed = flips)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flips: Internal") +
  ylab("Flips: gnomAD") +
  ggtitle("Flips: EAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_eagle_gnomad %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_eagle_int %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(TOPMed = p_flip)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flip rate: Internal") +
  ylab("Flip rate: gnomAD") +
  ggtitle("Flips / Number of Heterozygous Sites: EAGLE") +
  geom_abline(slope = 1, intercept = 0)
```

```{r}
df_wh_eagle_gnomad %>%
  select(id, pop, switches) %>%
  rename(gnomAD = switches) %>%
  inner_join({df_wh_eagle_int %>%
      select(id, pop, switches) %>%
      rename(TOPMed = switches)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switches: Internal") +
  ylab("Switches: gnomAD") +
  ggtitle("Switches: EAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_eagle_gnomad %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_eagle_int %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(TOPMed = p_flip)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switch rate: Internal") +
  ylab("Switch rate: gnomAD") +
  ggtitle("Switches / Number of Heterozygous Sites: EAGLE") +
  geom_abline(slope = 1, intercept = 0)
```

#### gnomAD and NYGC

```{r}
df_wh_eagle_gnomad %>%
  select(id, pop, flips) %>%
  rename(gnomAD = flips) %>%
  inner_join({df_wh_eagle_nygc %>%
      select(id, pop, flips) %>%
      rename(NYGC = flips)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flips: NYGC") +
  ylab("Flips: gnomAD") +
  ggtitle("Flips: EAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_eagle_gnomad %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_eagle_nygc %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flip rate: NYGC") +
  ylab("Flip rate: gnomAD") +
  ggtitle("Flips / Number of Heterozygous Sites: EAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_eagle_gnomad %>%
  select(id, pop, switches) %>%
  rename(gnomAD = switches) %>%
  inner_join({df_wh_eagle_nygc %>%
      select(id, pop, switches) %>%
      rename(NYGC = switches)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switches: NYGC") +
  ylab("Switches: gnomAD") +
  ggtitle("Switches: EAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_eagle_gnomad %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_eagle_nygc %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switch rate: NYGC") +
  ylab("Switch rate: gnomAD") +
  ggtitle("Switches / Number of Heterozygous Sites: EAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

#### TOPMed and NYGC

```{r}
df_wh_eagle_int %>%
  select(id, pop, flips) %>%
  rename(TOPMed = flips) %>%
  inner_join({df_wh_eagle_nygc %>%
      select(id, pop, flips) %>%
      rename(NYGC = flips)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Flips: NYGC") +
  ylab("Flips: TOPMed") +
  ggtitle("Flips: EAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_eagle_int %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(TOPMed = p_flip) %>%
  inner_join({df_wh_eagle_nygc %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Flip rate: NYGC") +
  ylab("Flip rate: TOPMed") +
  ggtitle("Flips / Number of Heterozygous Sites: EAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_eagle_int %>%
  select(id, pop, switches) %>%
  rename(TOPMed = switches) %>%
  inner_join({df_wh_eagle_nygc %>%
      select(id, pop, switches) %>%
      rename(NYGC = switches)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Switches: NYGC") +
  ylab("Switches: TOPMed") +
  ggtitle("Switches: EAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_eagle_int %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(TOPMed = p_flip) %>%
  inner_join({df_wh_eagle_nygc %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Switch rate: NYGC") +
  ylab("Switch rate: TOPMed") +
  ggtitle("Switches / Number of Heterozygous Sites: EAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

### BEAGLE

#### gnomAD and TOPMed

```{r}
df_wh_beagle_gnomad %>%
  select(id, pop, flips) %>%
  rename(gnomAD = flips) %>%
  inner_join({df_wh_beagle_int %>%
      select(id, pop, flips) %>%
      rename(TOPMed = flips)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flips: Internal") +
  ylab("Flips: gnomAD") +
  ggtitle("Flips: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_beagle_gnomad %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_beagle_int %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(TOPMed = p_flip)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flip rate: Internal") +
  ylab("Flip rate: gnomAD") +
  ggtitle("Flips / Number of Heterozygous Sites: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)
```

```{r}
df_wh_beagle_gnomad %>%
  select(id, pop, switches) %>%
  rename(gnomAD = switches) %>%
  inner_join({df_wh_beagle_int %>%
      select(id, pop, switches) %>%
      rename(TOPMed = switches)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switches: Internal") +
  ylab("Switches: gnomAD") +
  ggtitle("Switches: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_beagle_gnomad %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_beagle_int %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(TOPMed = p_flip)}) %>%
 ggplot(aes(x = TOPMed, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switch rate: Internal") +
  ylab("Switch rate: gnomAD") +
  ggtitle("Switches / Number of Heterozygous Sites: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)
```

#### gnomAD and NYGC

```{r}
df_wh_beagle_gnomad %>%
  select(id, pop, flips) %>%
  rename(gnomAD = flips) %>%
  inner_join({df_wh_beagle_nygc %>%
      select(id, pop, flips) %>%
      rename(NYGC = flips)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flips: NYGC") +
  ylab("Flips: gnomAD") +
  ggtitle("Flips: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_beagle_gnomad %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_beagle_nygc %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Flip rate: NYGC") +
  ylab("Flip rate: gnomAD") +
  ggtitle("Flips / Number of Heterozygous Sites: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_beagle_gnomad %>%
  select(id, pop, switches) %>%
  rename(gnomAD = switches) %>%
  inner_join({df_wh_beagle_nygc %>%
      select(id, pop, switches) %>%
      rename(NYGC = switches)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switches: NYGC") +
  ylab("Switches: gnomAD") +
  ggtitle("Switches: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_beagle_gnomad %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(gnomAD = p_flip) %>%
  inner_join({df_wh_beagle_nygc %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = gnomAD, colour = pop)) +
  geom_point() +
  xlab("Switch rate: NYGC") +
  ylab("Switch rate: gnomAD") +
  ggtitle("Switches / Number of Heterozygous Sites: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

#### TOPMed and NYGC

```{r}
df_wh_beagle_int %>%
  select(id, pop, flips) %>%
  rename(TOPMed = flips) %>%
  inner_join({df_wh_beagle_nygc %>%
      select(id, pop, flips) %>%
      rename(NYGC = flips)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Flips: NYGC") +
  ylab("Flips: TOPMed") +
  ggtitle("Flips: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_beagle_int %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(TOPMed = p_flip) %>%
  inner_join({df_wh_beagle_nygc %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Flip rate: NYGC") +
  ylab("Flip rate: TOPMed") +
  ggtitle("Flips / Number of Heterozygous Sites: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_beagle_int %>%
  select(id, pop, switches) %>%
  rename(TOPMed = switches) %>%
  inner_join({df_wh_beagle_nygc %>%
      select(id, pop, switches) %>%
      rename(NYGC = switches)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Switches: NYGC") +
  ylab("Switches: TOPMed") +
  ggtitle("Switches: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_beagle_int %>%
  mutate(p_flip = switches / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(TOPMed = p_flip) %>%
  inner_join({df_wh_beagle_nygc %>%
      mutate(p_flip = switches / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(NYGC = p_flip)}) %>%
 ggplot(aes(x = NYGC, y = TOPMed, colour = pop)) +
  geom_point() +
  xlab("Switch rate: NYGC") +
  ylab("Switch rate: TOPMed") +
  ggtitle("Switches / Number of Heterozygous Sites: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```
