---
title: "09012023_consensus_idea"
author: "Andy Beck"
date: "2023-01-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

The idea here is to take the phased VCFs from EAGLE, BEAGLE, and SHAPEIT and somehow combine them in order to come up with a single pair of haploids.

```{r}
library(tidyverse)
id_df <- read_csv("/net/snowwhite/home/beckandy/research/phasing/data/sample_pairs_9aug2023.csv", 
                  col_names = c("POP", "ID1", "ID2"))
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
library(ggsci)
```

```{r}
whats_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/whatshap/"

# df_wh_eagle <- read_tsv(paste0(whats_dir, "eagle/eval_1.tsv"))
# df_wh_beagle <- read_tsv(paste0(whats_dir, "beagle/eval_1.tsv"))
# df_wh_shapeit <- read_tsv(paste0(whats_dir, "shapeit/eval_1.tsv"))
# 
# df_wh_eagle$all_switches
# df_wh_beagle$all_switches
# df_wh_shapeit$all_switches
# 
# df_wh_eagle$all_switchflips
# df_wh_beagle$all_switchflips
# df_wh_shapeit$all_switchflips
```

# Result Analysis

Having run the voting process on all 700 pseudo-diploids, let's load in the results for all samples.

```{r}
get_vote_res <- function(ids = 700, 
                         res_dir = "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/vote/",
                         prefix = "error_vote_",
                         suffix = ".tsv"){
  df <- read_tsv(paste0(res_dir, prefix, 1, suffix), show_col_types = FALSE) %>% 
    select(starts_with("all"))
  for(i in 2:ids){
    df <- bind_rows(df,
                    {read_tsv(paste0(res_dir, prefix, i, suffix), show_col_types = FALSE) %>%
                        select(starts_with("all"))})
  }
  df$id <- 1:ids
  return(df)
}

# df <- get_vote_res()
# df_easy <- get_vote_res(res_dir = "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/vote/easy/")
# df$id <- 1:700
# df_easy$id <- 1:700
# df$alg <- "vote"
# df_easy$alg <-"simple_vote"
```

Let's also gather these results for the 3 phasing algorithms:

```{r}
get_alg_res <- function(alg, 
                        data_dir = "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/whatshap/",
                        ids = 700){
  base_dir <- paste0(data_dir, alg, "/")
  df <- read_tsv(paste0(base_dir, "eval_", 1,".tsv"), show_col_types = FALSE) %>%
    select(starts_with("all"))
  for(i in 2:ids){
    df <- bind_rows(df, {
      read_tsv(paste0(base_dir, "eval_", i,".tsv"), show_col_types = FALSE) %>%
        select(starts_with("all"))
    })
  }
  return(df)
}
# df_beagle <- get_alg_res("beagle")
# df_eagle <- get_alg_res("eagle")
# df_shapeit <- get_alg_res("shapeit")
# 
# df_beagle$id <- 1:700
# df_eagle$id <- 1:700
# df_shapeit$id <- 1:700
# 
# df_beagle$alg <- "beagle"
# df_eagle$alg <- "eagle"
# df_shapeit$alg <- "shapeit"

# smash everything together
# df <- bind_rows(df, df_beagle) %>%
#   bind_rows(df_eagle) %>%
#   bind_rows(df_shapeit) %>%
#   bind_rows(df_easy)

# save for later usage
# write_csv(df, "/net/snowwhite/home/beckandy/research/phasing/data/consensus/whatshap_counts.csv")
```

```{r}
df <- read_csv("/net/snowwhite/home/beckandy/research/phasing/data/consensus/whatshap_counts.csv")

df$pop <- c(
  rep(
    c(rep(c("EUR", "AFR"), each = 200), rep(c("AMR", "EAS", "SAS"), each = 100))
    , times = 5))
```

Plots by algorithm

```{r}
df %>%
  select(id, pop, alg, all_switches) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = all_switches) %>%
  ggplot(aes(x = simple_vote, y = vote, colour = pop)) +
  geom_point() +
  xlab("EASY") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Total Errors", "EASY and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))

df %>%
  select(id, pop, alg, all_switches) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = all_switches) %>%
  ggplot(aes(x = beagle, y = vote, colour = pop)) +
  geom_point() +
  xlab("BEAGLE") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Total Errors", "BEAGLE and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))

df %>%
  select(id, pop, alg, all_switches) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = all_switches) %>%
  ggplot(aes(x = eagle, y = vote, colour = pop)) +
  geom_point() +
  xlab("EAGLE") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Total Errors", "EAGLE and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))

df %>%
  select(id, pop, alg, all_switches) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = all_switches) %>%
  ggplot(aes(x = shapeit, y = vote, colour = pop)) +
  geom_point() +
  xlab("SHAPEIT") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Total Errors", "SHAPEIT and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))
```

Separate flips from switches:

```{r}
df <- df %>%
  separate(all_switchflips, into = c("switches", "flips"), sep = "/", remove = FALSE) %>%
  mutate(switches = as.numeric(switches),
         flips = as.numeric(flips))
```

Just switches:

```{r}
df %>%
  select(id, pop, alg, switches) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = switches) %>%
  ggplot(aes(x = beagle, y = vote, colour = pop)) +
  geom_point() +
  xlab("BEAGLE") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Switches", "BEAGLE and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))

df %>%
  select(id, pop, alg, switches) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = switches) %>%
  ggplot(aes(x = eagle, y = vote, colour = pop)) +
  geom_point() +
  xlab("EAGLE") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Switches", "EAGLE and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))

df %>%
  select(id, pop, alg, switches) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = switches) %>%
  ggplot(aes(x = shapeit, y = vote, colour = pop)) +
  geom_point() +
  xlab("SHAPEIT") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Switches", "SHAPEIT and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))
```

```{r}
df %>%
  select(id, pop, alg, flips) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = flips) %>%
  ggplot(aes(x = beagle, y = vote, colour = pop)) +
  geom_point() +
  xlab("BEAGLE") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Flips", "BEAGLE and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))

df %>%
  select(id, pop, alg, flips) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = flips) %>%
  ggplot(aes(x = eagle, y = vote, colour = pop)) +
  geom_point() +
  xlab("EAGLE") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Flips", "EAGLE and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))

df %>%
  select(id, pop, alg, flips) %>%
  pivot_wider(id_cols = c(id, pop), names_from = alg, values_from = flips) %>%
  ggplot(aes(x = shapeit, y = vote, colour = pop)) +
  geom_point() +
  xlab("SHAPEIT") +
  ylab("VOTE") +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Flips", "SHAPEIT and VOTE") +
  scale_color_jco() +
  theme_classic(base_size = 13) +
  theme(legend.position = c(0.8, .2))
```
