---
title: "16nov22_more_cpg"
author: "Andy Beck"
date: "2022-11-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


## Libraries and global variables

```{r}
library(tidyverse)
library(janitor)
source("code/common_functions.R")

library(ggsci)
theme_set(theme_bw(base_size = 18))

pal1 <- pal_ucscgb()

eagle_switch_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/switch_errors/eagle/annotated/"
shapeit_switch_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/switch_errors/shapeit/annotated/"
beagle_switch_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/switch_errors/beagle/annotated/"
num_sites_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/27sept/vcf_n_sites/"
whatshap_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/whatshap/"
het_loc_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/het_loc/annotated/"

pair_info_df <- read_delim("data/sample_pairs_16aug2022.csv", col_names = c("POP", "ID1", "ID2"))
pair_info_df$SP <- c(rep("EUR", 200), rep("AFR", 200))

gc_content_1kb <- read_tsv("data/gc1kb_X_only.bed")
colnames(gc_content_1kb) <- c("CHR", "START", "END", "AT", "GC", "A", "C", "G", "T", "TOTAL", "OTHER", "LENGTH")
gc_content_1kb  <- gc_content_1kb %>%
  mutate(bin_id = (START / 1000) + 1)

background_dimer <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/background_rates/dimer.csv")
background_3mer <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/background_rates/3mer.csv")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

Here in this document we are going to take a more "refined" look at the enrichment of errors at CpG sites. Namely, instead of using the rate of CpG dimers along the entire chromosome as our baseline, we
instead look at the proportion of heterozygous sites within each pseudo-diploid that are CpGs as 
our baseline benchmark (with the assumption going in that this won't impact our analyses much, but it's 
"more correct" or whatever)

```{r}
df_vcftools <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, eagle_switch_dir, beagle_switch_dir, shapeit_switch_dir, gc_content_1kb)
                      }) %>%
  bind_rows()

df_vcftools$pop <- c(rep("EUR", 200), rep("AFR", 200))
```

```{r}
df_wh_eagle <- get_all_whatshap("eagle")
df_wh_beagle <- get_all_whatshap("beagle")
df_wh_si <- get_all_whatshap("shapeit")
```

Now let's append proportion heterozygous CpG info to each pair

```{r}
df_hets <- data.frame(pair_id = 1:400, nhet = numeric(400), cpg_het = numeric(400))

for(i in 1:400){
  f_name <- paste0(het_loc_dir, "pair_", i, "_het_loc.txt")
  tmp <- read_tsv(f_name, show_col_types = FALSE)
  df_hets$nhet[i] <- length(tmp$cpg)
  df_hets$cpg_het[i] <- sum(tmp$cpg)
  rm(tmp)
}

df_hets$prop_cpg <- df_hets$cpg_het / df_hets$nhet

df_hets %>%
  ggplot(aes(x = nhet, y = cpg_het))+
  geom_point() +
  ggtitle("CpG Heterozygous Sites") +
  xlab("Total Heterozygous Sites") +
  ylab("CpG Heterozygous Sites")

df_hets %>%
  ggplot(aes(x = prop_cpg))+
  geom_density() +
  ggtitle("Proportion CpG Heterozygous Sites")
```

```{r}
df_vcftools %>%
  ggplot(aes(y = prop_cpg)) + geom_boxplot()+ geom_hline(yintercept = background_cpg_het)
```


```{r}
df_vcftools <- inner_join(df_vcftools, df_hets, by = "pair_id")

df_vcftools %>%
  mutate(prop_error_eagle = n_switch_cpg_eagle / n_switch_eagle,
         prop_error_shapeit = n_switch_cpg_shapeit / n_switch_shapeit,
         prop_error_beagle = n_switch_cpg_beagle / n_switch_beagle) %>%
  mutate(enrich_eagle = prop_error_eagle / prop_cpg,
         enrich_beagle = prop_error_beagle / prop_cpg,
         enrich_shapeit = prop_error_shapeit / prop_cpg) %>%
  select(pop, starts_with("enrich")) %>%
  rename(EAGLE2 = enrich_eagle,
         BEAGLE5 = enrich_beagle,
         SHAPEIT4 = enrich_shapeit) %>%
  pivot_longer(-pop, names_to = "method",  values_to = "prop") %>%
  ggplot(aes(x = method, y = prop)) + 
  geom_boxplot() +
  xlab("Phasing Method") +
  ylab("Enrichment of errors at CpG") +
  scale_color_manual(values = "red") +
  ggtitle("Enrichment of Phasing Errors at CpG Sites") +
  theme_bw(base_size = 16) + labs(color = "")
```

$$
\Pr(Error | CpG) = \frac{\Pr(CpG | Error)\Pr(error)}{\Pr(CpG)}
$$

```{r}
df_vcftools %>%
  mutate(p_error_eagle = n_switch_eagle / nhet,
         p_error_beagle = n_switch_beagle / nhet,
         p_error_shapeit = n_switch_shapeit / nhet) %>%
  mutate(prop_error_eagle = n_switch_cpg_eagle / n_switch_eagle,
         prop_error_shapeit = n_switch_cpg_shapeit / n_switch_shapeit,
         prop_error_beagle = n_switch_cpg_beagle / n_switch_beagle) %>%
  mutate(p_cpg_eagle = prop_error_eagle * p_error_eagle / prop_cpg) %>%
  ggplot(aes(x = p_cpg_eagle)) + geom_density()
  
```

