---
title: "20220131_compare_filter_non"
author: "Andy Beck"
date: "2023-01-31"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(tidyverse)
library(janitor)
source('code/common_functions.R')

pair_info_df <- read_delim("data/sample_pairs_16aug2022.csv", col_names = c("POP", "ID1", "ID2"))
pair_info_df$SP <- c(rep("EUR", 200), rep("AFR", 200))

gc_content_1kb <- read_tsv("data/gc1kb_X_only.bed")
colnames(gc_content_1kb) <- c("CHR", "START", "END", "AT", "GC", "A", "C", "G", "T", "TOTAL", "OTHER", "LENGTH")
gc_content_1kb  <- gc_content_1kb %>%
  mutate(bin_id = (START / 1000) + 1)
```

In this document we will be comparing the rates of errors from two different runs of constructing our pseudo-diploids. In the first my understanding is that singletons might not have been properly removed from the vcf since I was using bcftool's `-c 2` option, which would not exclude a singleton male heterozygote since it would have been coded `1/1` and thus AC >= 2. The latter run uses a vcf in which sites from `vcftools --singletons` were removed.

Let's first load the vcftools results:

```{r}
base_dir_old <- "/net/snowwhite/home/beckandy/research/phasing/output/2023_switch_errors/switch_errors/"
eagle_dir_old <- paste0(base_dir_old, "eagle/annotated/")
bagle_dir_old <- paste0(base_dir_old, "beagle/annotated/")
shapeit_dir_old <- paste0(base_dir_old, "shapeit/annotated/")

df_vcftools_old <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, eagle_dir_old, bagle_dir_old, shapeit_dir_old, gc_content_1kb)
                      }) %>%
  bind_rows()
df_vcftools_old$pop <- c(rep("EUR", 200), rep("AFR", 200))

base_dir_new <- "/net/snowwhite/home/beckandy/research/phasing/output/2023_switch_errors/switch_errors/"
eagle_dir_new <- paste0(base_dir_new, "eagle/annotated/")
bagle_dir_new <- paste0(base_dir_new, "beagle/annotated/")
shapeit_dir_new <- paste0(base_dir_new, "shapeit/annotated/")

df_vcftools_new <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, eagle_dir_new, bagle_dir_new, shapeit_dir_new, gc_content_1kb)
                      }) %>%
  bind_rows()
df_vcftools_new$pop <- c(rep("EUR", 200), rep("AFR", 200))
```

And now from whatshap:

```{r}
whatshap_dir_old <- "/net/snowwhite/home/beckandy/research/phasing/output/2023_switch_errors/whatshap"
df_wh_eagle_old <- get_all_whatshap("eagle", whatshap_dir_old)
df_wh_beagle_old <- get_all_whatshap("beagle", whatshap_dir_old)
df_wh_si_old <- get_all_whatshap("shapeit", whatshap_dir_old)

whatshap_dir_new <- "/net/snowwhite/home/beckandy/research/phasing/output/filter_switch_errors/whatshap"
df_wh_eagle_new <- get_all_whatshap("eagle", whatshap_dir_new)
df_wh_beagle_new <- get_all_whatshap("beagle", whatshap_dir_new)
df_wh_si_new <- get_all_whatshap("shapeit", whatshap_dir_new)
```

## Flips and Swtiches Comparison

### SHAPEIT

```{r}
df_wh_si_new %>%
  select(id, pop, flips) %>%
  rename(new = flips) %>%
  inner_join({df_wh_si_old %>%
      select(id, pop, flips) %>%
      rename(old = flips)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Flips previous") +
  ylab("Flips with filter") +
  ggtitle("Flips: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)

df_wh_si_new %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(new = p_flip) %>%
  inner_join({df_wh_si_old %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(old = p_flip)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Flip rate previous") +
  ylab("Flip rate with filter") +
  ggtitle("Flips / Number of Heterozygous Sites: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_si_new %>%
  select(id, pop, switches) %>%
  rename(new = switches) %>%
  inner_join({df_wh_si_old %>%
      select(id, pop, switches) %>%
      rename(old = switches)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Switches previous") +
  ylab("Switches with filter") +
  ggtitle("Switches: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)

df_wh_si_new %>%
  mutate(p_switch = switches / n_het) %>%
  select(id, pop, p_switch) %>%
  rename(new = p_switch) %>%
  inner_join({df_wh_si_old %>%
      mutate(p_switch = switches / n_het) %>%
      select(id, pop, p_switch) %>%
      rename(old = p_switch)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Switch rate previous") +
  ylab("Switch rate with filter") +
  ggtitle("Switch / Number of Heterozygous Sites: SHAPEIT") +
  geom_abline(slope = 1, intercept = 0)  
```

### BEAGLE

```{r}
df_wh_beagle_new %>%
  select(id, pop, flips) %>%
  rename(new = flips) %>%
  inner_join({df_wh_beagle_old %>%
      select(id, pop, flips) %>%
      rename(old = flips)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Flips previous") +
  ylab("Flips with filter") +
  ggtitle("Flips: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_beagle_new %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(new = p_flip) %>%
  inner_join({df_wh_beagle_old %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(old = p_flip)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Flip rate previous") +
  ylab("Flip rate with filter") +
  ggtitle("Flips / Number of Heterozygous Sites: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_beagle_new %>%
  select(id, pop, switches) %>%
  rename(new = switches) %>%
  inner_join({df_wh_beagle_old %>%
      select(id, pop, switches) %>%
      rename(old = switches)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Switches previous") +
  ylab("Switches with filter") +
  ggtitle("Switches: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_beagle_new %>%
  mutate(p_switch = switches / n_het) %>%
  select(id, pop, p_switch) %>%
  rename(new = p_switch) %>%
  inner_join({df_wh_beagle_old %>%
      mutate(p_switch = switches / n_het) %>%
      select(id, pop, p_switch) %>%
      rename(old = p_switch)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Switch rate previous") +
  ylab("Switch rate with filter") +
  ggtitle("Switch / Number of Heterozygous Sites: BEAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

### EAGLE

```{r}
df_wh_eagle_new %>%
  select(id, pop, flips) %>%
  rename(new = flips) %>%
  inner_join({df_wh_eagle_old %>%
      select(id, pop, flips) %>%
      rename(old = flips)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Flips previous") +
  ylab("Flips with filter") +
  ggtitle("Flips: EAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_eagle_new %>%
  mutate(p_flip = flips / n_het) %>%
  select(id, pop, p_flip) %>%
  rename(new = p_flip) %>%
  inner_join({df_wh_eagle_old %>%
      mutate(p_flip = flips / n_het) %>%
      select(id, pop, p_flip) %>%
      rename(old = p_flip)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Flip rate previous") +
  ylab("Flip rate with filter") +
  ggtitle("Flips / Number of Heterozygous Sites: EAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

```{r}
df_wh_eagle_new %>%
  select(id, pop, switches) %>%
  rename(new = switches) %>%
  inner_join({df_wh_eagle_old %>%
      select(id, pop, switches) %>%
      rename(old = switches)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Switches previous") +
  ylab("Switches with filter") +
  ggtitle("Switches: EAGLE") +
  geom_abline(slope = 1, intercept = 0)

df_wh_eagle_new %>%
  mutate(p_switch = switches / n_het) %>%
  select(id, pop, p_switch) %>%
  rename(new = p_switch) %>%
  inner_join({df_wh_eagle_old %>%
      mutate(p_switch = switches / n_het) %>%
      select(id, pop, p_switch) %>%
      rename(old = p_switch)}) %>%
 ggplot(aes(x = old, y = new, colour = pop)) +
  geom_point() +
  xlab("Switch rate previous") +
  ylab("Switch rate with filter") +
  ggtitle("Switch / Number of Heterozygous Sites: EAGLE") +
  geom_abline(slope = 1, intercept = 0)  
```

