---
title: "gnomAD Data"
author: "Andy Beck"
date: "2023-02-21"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r}
library(tidyverse)
```

## Introduction

The purpose of this document is to catalog data preparation and analysis using the gnomAD v3 release of individual-level genotype data for the 
1KGP + HDP data sets. 


## Sample Metadata

Downloaded from [gnomAD's download page](https://gnomad.broadinstitute.org/downloads#v3). Tab-delimited files with multiple variables in several columns which we'll need to parse (yay)

```{r}
df_meta <- read_tsv("data/gnomAD/gnomad.genomes.v3.1.2.hgdp_1kg_subset_sample_meta.tsv")

df_meta %>%
  select(s, hgdp_tgp_meta ) %>%
  head()
```

It looks as though the information we want initially is buried in the `hgdp_tgp_meta` column. Let's see if we can wrangle this into separate columns:

```{r}
# Remove quotes and curly brackets
df_meta <- df_meta %>%
  mutate(info = str_remove_all(hgdp_tgp_meta, "[{}\"]"))

# get first four columns of this column
df_study_info <- df_meta %>%
  rowwise() %>%
  mutate(project = str_split(info, ",")[[1]][1],
         study_region = str_split(info, ",")[[1]][2],
         population = str_split(info, ",")[[1]][3],
         genetic_region = str_split(info, ",")[[1]][4]) %>%
  mutate(project = str_split(project, ":")[[1]][2],
         study_region = str_split(study_region, ":")[[1]][2],
         population = str_split(population, ":")[[1]][2],
         genetic_region = str_split(genetic_region, ":")[[1]][2]) %>%
  select(s, project, study_region, population, genetic_region)

df_meta <-  df_meta %>%
  rowwise() %>%
  mutate(project = str_split(info, ",")[[1]][1],
         study_region = str_split(info, ",")[[1]][2],
         population = str_split(info, ",")[[1]][3],
         genetic_region = str_split(info, ",")[[1]][4]) %>%
  mutate(project = str_split(project, ":")[[1]][2],
         study_region = str_split(study_region, ":")[[1]][2],
         population = str_split(population, ":")[[1]][2],
         genetic_region = str_split(genetic_region, ":")[[1]][2])
```

Let's see how many samples were included in gnomAD release:

```{r}
df_meta %>%
  filter(gnomad_release == TRUE) %>%
  pull(project) %>%
  table()
```

Next, out of the 2500 released 1KGP samples, how many are in our list of unrelated samples?

```{r}
unrel_id <- read_tsv("data/1kgp/unrelated_subj.tsv") %>% pull(SAMPLE_NAME)

keep_id <- intersect({df_meta %>% filter(gnomad_release == TRUE) %>% pull(s)}, unrel_id)
```

Write the names of the 1KGP samples into a file:

```{r}
# df_meta %>%
#   filter(gnomad_release == TRUE) %>%
#   filter(project == "1000 Genomes") %>%
#   select(s) %>%
#   write_csv("/net/snowwhite/home/beckandy/research/phasing/data/gnomAD/1kgp_samples.txt", col_names = FALSE, quote = "none")

data.frame(id = keep_id) %>% write_csv("/net/snowwhite/home/beckandy/research/phasing/data/gnomAD/1kgp_samples.txt", col_names = FALSE, quote = "none")
```

