---
title: "Parameter Test"
author: "Andy Beck"
date: "2022-09-28"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(janitor)
source("code/common_functions.R")

# Overwrite the switch_summary function
switch_summary <- function(pair_id, input_dir, gc_content_1kb, bin_size = 1000){
  switch_err_df <- read_csv(paste0(input_dir, "switch_", pair_id, ".csv"), show_col_types = FALSE) %>%
    mutate(bin_id = ceiling(pos_start / bin_size))
  
  switch_err_df <- switch_err_df %>%
    left_join({gc_content_1kb %>% select(bin_id, GC)}, by = "bin_id")
  
  # get positions of flips
  flip_pos <- get_flip_pos(switch_err_df)
  
  # Assign switches flip status
  switch_err_df$is_flip <- (switch_err_df$pos_start %in% flip_pos) |
    (switch_err_df$pos_end %in% flip_pos)
  
  # Start of flip
  switch_err_df$start_flip <- (switch_err_df$pos_start %in% flip_pos)
  
  # stats we want to pull
  n_switch <- length(switch_err_df$pos_start)
  n_flip <- sum(switch_err_df$is_flip) / 2
  n_other <- n_switch - 2*n_flip

  n_switch_cpg <- switch_err_df %>%
    filter(cpg_start == 1) %>%
    pull(is_flip) %>%
    length()

  n_flip_cpg <- switch_err_df %>%
    filter(cpg_start == 1) %>%
    pull(start_flip) %>%
    sum()
  
  mean_gc_switch <- mean(switch_err_df$GC)
  
  return(data.frame(pair_id = pair_id,
                    n_switch = n_switch,
                    n_flip = n_flip,
                    n_other = n_other,
                    n_switch_cpg = n_switch_cpg,
                    n_flip_cpg = n_flip_cpg,
                    mean_gc_switch = mean_gc_switch
                    ))
}

pair_info_df <- read_delim("data/sample_pairs_16aug2022.csv", col_names = c("POP", "ID1", "ID2"))
pair_info_df$SP <- c(rep("EUR", 200), rep("AFR", 200))
pair_info_df$pair_id <- 1:400

gc_content_1kb <- read_tsv("data/gc1kb_X_only.bed")
colnames(gc_content_1kb) <- c("CHR", "START", "END", "AT", "GC", "A", "C", "G", "T", "TOTAL", "OTHER", "LENGTH")
gc_content_1kb  <- gc_content_1kb %>%
  mutate(bin_id = (START / 1000) + 1)

background_dimer <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/background_rates/dimer.csv")
background_3mer <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/background_rates/3mer.csv")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

background_cpg_het <- 229868 / 1567042
```

## Introduction

The purpose of this document is to explore the effects of changing parameters for the phasing algorithms. The impetus for this was two observations:

1. SHAPEIT4 has a parameter that changes some defaults when phasing data from sequencing (`--phasing`) which I was unaware of in my first round of simulations
2. BEAGLE appeared to have done better when no genetic map was included.

This document will focus on the above two issues first, and later we may look more thoroughly into other parameter choices which may improve the phasing of our pseudo-diploids.

## First Analysis

### SHAPEIT

Let's first take a look at the SHAPEIT4 results: does the `--sequencing` option reduce the number of errors?

```{r}
shapeit_np_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/switch_errors/shapeit/annotated/"
shapeit_p_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/27sept/switch_errors/shapeit/annotated/"

whatshap_np <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/whatshap/"
whatshap_p <- "/net/snowwhite/home/beckandy/research/phasing/output/27sept/whatshap/"

df_s_np <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, shapeit_np_dir, gc_content_1kb)
                      }) %>%
  bind_rows()

df_s_p <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, shapeit_p_dir, gc_content_1kb)
                      }) %>%
  bind_rows()

df_s_np$type <- "no_sequencing"
df_s_p$type <- "sequencing"

df <- bind_rows(df_s_np, df_s_p)
```

Let's first look at the number of total errors we see in each pseudo-diploid:

```{r}
df %>%
  select(pair_id, n_switch, type) %>%
  pivot_wider(id_cols = pair_id, names_from = type, values_from = n_switch) %>%
  left_join(pair_info_df, by = "pair_id") %>%
  ggplot(aes(x = no_sequencing, y = sequencing, color = SP)) +
  geom_point() +
  ggtitle("Total Errors: SHAPEIT4", "With and without --sequencing option") +
  xlab("With sequencing parameters") + ylab("Without") +
  geom_abline(slope = 1, intercept = 0)
```

Not much of a difference here; let's take a look at flips:

```{r}
df %>%
  select(pair_id, n_flip, type) %>%
  pivot_wider(id_cols = pair_id, names_from = type, values_from = n_flip) %>%
  left_join(pair_info_df, by = "pair_id") %>%
  ggplot(aes(x = no_sequencing, y = sequencing, color = SP)) +
  geom_point() +
  ggtitle("Flips: SHAPEIT4", "With and without --sequencing option") +
  xlab("With sequencing parameters") + ylab("Without") +
  geom_abline(slope = 1, intercept = 0)
```

### BEAGLE

Let's take another look at BEAGLE with and without the genetic map.

```{r}
beagle_gm_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/switch_errors/beagle/annotated/"
beagle_nm_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/27sept/switch_errors/beagle/annotated/"

whatshap_gm <- "/net/snowwhite/home/beckandy/research/phasing/output/9aug22_switch_errors/whatshap/"
whatshap_nm <- "/net/snowwhite/home/beckandy/research/phasing/output/27sept/whatshap/"

df_gm <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, beagle_gm_dir, gc_content_1kb)
                      }) %>%
  bind_rows()

df_nm <- lapply(c(1:400), 
                    function(x){
                      switch_summary(x, beagle_nm_dir, gc_content_1kb)
                      }) %>%
  bind_rows()

df_gm$type <- "genetic_map"
df_nm$type <- "constant_recomb"

df <- bind_rows(df_gm, df_nm)
```

Let's first look at the number of errors

```{r}
df %>%
  select(pair_id, n_switch, type) %>%
  pivot_wider(id_cols = pair_id, names_from = type, values_from = n_switch) %>%
  left_join(pair_info_df, by = "pair_id") %>%
  ggplot(aes(x = genetic_map, y = constant_recomb, color = SP)) +
  geom_point() +
  ggtitle("Total Errors: BEAGLE 5.2", "Genetic Map versus Constant Recombination Rate") +
  xlab("With genetic map") + ylab("Constant Rate") +
  geom_abline(slope = 1, intercept = 0)
```

and now flips:

```{r}
df %>%
  select(pair_id, n_flip, type) %>%
  pivot_wider(id_cols = pair_id, names_from = type, values_from = n_flip) %>%
  left_join(pair_info_df, by = "pair_id") %>%
  ggplot(aes(x = genetic_map, y = constant_recomb, color = SP)) +
  geom_point() +
  ggtitle("Flips: BEAGLE 5.2", "Genetic Map versus Constant Recombination Rate") +
  xlab("With genetic map") + ylab("Constant Rate") +
  geom_abline(slope = 1, intercept = 0)
```

