---
title: "gc_content"
author: "Andy Beck"
date: "2022-04-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r}
library(tidyverse)
```


## Introduction

In this document we'll take a peek at what the gc content looks like genome-wide. For now we have this available in windows of size 10kb and 100kb (it might be worth asking if I should be looking at larger windows). Let's first look at the 10kb windows

```{r}
col_names <- c("CHR", "START", "END", "AT", "GC", "A", "C", "G", "T", "N", "OTHER", "LEN")
df_10k <- read_tsv("data/gc10kb.bed")
df_100k <- read_tsv("data/gc100kb.bed")
df_1k <- read_tsv("data/gc1kb_X_only.bed")

colnames(df_1k) <- col_names
colnames(df_10k) <- col_names
colnames(df_100k) <- col_names
```


Let's take a peek at the X chromosome:

```{r}
for(chrom in c("X")){
  c_name <- paste0("chr", chrom)
  p<- df_10k %>%
    filter(CHR == c_name) %>%
    ggplot(aes(x = START, y = GC)) + 
    geom_line() + 
    geom_smooth() +
    ggtitle("% GC Content in 10kb Windows", paste0("Chromosome ", chrom)) +
    xlab("Window Start Position") +
    ylab("% GC")
  print(p)
}
```

```{r}
for(chrom in c("X")){
  c_name <- paste0("chr", chrom)
  p<- df_100k %>%
    filter(CHR == c_name) %>%
    ggplot(aes(x = START, y = GC)) + 
    geom_line() + 
    geom_smooth() +
    ggtitle("% GC Content in 100kb Windows", paste0("Chromosome ", chrom)) +
    xlab("Window Start Position") +
    ylab("% GC")
  print(p)
}
```

```{r}
df_1k %>%
  ggplot(aes(x = START, y = GC))+
  geom_line() +
  geom_smooth()+
  ggtitle("% GC Content in 1kb Windows", "Chromosome X") +
  xlab("Window Start Position") +
  ylab("% GC")
```

## Assigning bins to switches

```{r}
df_1k <- df_1k %>%
  rowwise() %>%
  mutate(bin_id = (START / 1000) + 1)
df_1k %>% head
```

## Single pseudo-diploid Exploration

```{r}
df_eagle <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/18mar22_switch_errors/switch_errors/eagle/annotated/switch_1.csv")
df_shapeit <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/18mar22_switch_errors/switch_errors/shapeit/annotated/switch_1.csv")

background_dimer <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/background_rates/dimer.csv")
background_3mer <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/background_rates/3mer.csv")
```

Add bin ids for start (and end?):

```{r}
df_eagle <- df_eagle %>%
  mutate(start_bin = ceiling(pos_start / 1000),
         end_bin = ceiling(pos_end / 1000))

df_shapeit <- df_shapeit %>%
  mutate(start_bin = ceiling(pos_start / 1000),
         end_bin = ceiling(pos_end / 1000))
```

Match GC content:

```{r}
df_1k %>% 
  select(bin_id, GC) %>%
  right_join(df_shapeit, by = c("bin_id" = "start_bin")) %>%
  pull(GC) %>% mean()
```

