---
title: "methylation_explore"
author: "Andy Beck"
date: "2022-05-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


## Introduction

File: `data/methylation_chrX.bed`

```{r}
background_meth <- read_tsv("data/methylation_chrX.bed", 
                    col_names = c("chrom", "start", "stop", "meth"))

background_meth %>%
  head() %>%
  knitr::kable()
```

Let's plot these values by position:

```{r}
background_meth %>%
  ggplot(aes(x = start, y = meth)) +
  geom_point() +
  xlab("Start Position") +
  ylab("Measure")
```

Wow, very helpful; let's take a look at the distribution of the range of values for the measure:

```{r}
background_meth %>%
  ggplot(aes(x = meth)) + 
  geom_histogram()
```

Okay, so if I use the rule >50% = methylated, what does our distribution of methylated vs unmethylated sites look like?

```{r}

background_meth %>%
  mutate(is_meth = meth > 0.5) %>%
  tabyl(is_meth)

background_meth_rate <- background_meth %>%
  mutate(is_meth = meth > 0.5) %>%
  {sum(.$is_meth) / length(.$is_meth)}

```

```{r}
meth_df[sample(1:nrow(meth_df), 50000, replace = F),] %>%
  mutate(is_meth = as.numeric(meth > 0.5)) %>%
  ggplot(aes(x = start, y = is_meth, colour = factor(is_meth))) +
  geom_point()
```

Other things I might want to look at?

* Runs of methylation

## Runs of methylation

```{r}
meth_df <- meth_df %>%
  mutate(is_meth = as.numeric(meth > 0.5))

write_csv(meth_df, file = "output/meth_locations.csv")
```

```{r}
meth_runs <- function(df){
  final <- data.frame(start = numeric(), end = numeric(), type = numeric(), n = numeric())
  current_start <- df$start[1]
  current_end <- df$stop[1]
  current_type <- df$is_meth[1]
  counts <- 1
  
  for(i in 2:nrow(df)){
    if(df$is_meth[i] == current_type){
      current_end = df$stop[i]
      counts <- counts + 1
    } else{
      final <- bind_rows(final, data.frame(start = current_start, end = current_end, type = current_type, n = counts))
      current_start = df$start[i]
      current_end = df$stop[i]
      current_type = df$is_meth[i]
      counts <- 1
    }
  }
  final <- bind_rows(final, data.frame(start = current_start, end = current_end, type = current_type, n = counts))
  return(final)
}
```

```{r}
runny_meth <- meth_runs(meth_df)
```

```{r}
runny_meth %>%
  ggplot(aes(x = n, y = type, group_by = factor(type), colour = factor(type))) + 
  geom_boxplot()
```

```{r}
runny_meth %>%
  ggplot(aes(x = start, y = n, colour = factor(type))) +
  geom_point()
```

## Compare to Ovary

```{r}
# https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM1010980

ovary_df <- read_tsv("data/ovary/ovary_chrX.bed", 
                    col_names = c("chrom", "start", "stop", "meth"))

all_meth <- full_join(ovary_df, meth_df, by = c("chrom", "start", "stop")) %>%
  rename(ovary = meth.x, testis = meth.y) 

all_meth %>% summary()
```

```{r}
ovary_df %>%
  ggplot(aes(x = meth)) +
  geom_boxplot() +
  ggtitle("Methylation: Ovary (X Chromosome)")
```

```{r eval=FALSE}
all_meth %>%
  remove_missing() %>%
  ggplot(aes(x = testis, y = ovary)) + geom_point()
```

```{r}
all_meth %>%
  select(testis, ovary) %>%
  pivot_longer(testis:ovary, names_to = "sample", values_to = "methylation") %>%
  ggplot(aes(x = methylation, y= sample, colour = sample)) + geom_boxplot()
```

```{r}
all_meth %>%
  drop_na(ovary, testis) %>%
  mutate(ov_m = ovary > 0.5, ov_t = testis > 0.5) %>%
  select(ov_m, ov_t) %>%
  xtabs(~ ov_m + ov_t, data = .)
```



## Methylated Annotated Switches

```{r}
get_flip_pos <- function(df){
  flip_list <- c()
  in_progress <- FALSE
  for(i in 1:(length(df$pos_start)-1)){
    if(df$pos_end[i] == df$pos_start[i+1]){
      if(!in_progress){
        in_progress <- TRUE
        flip_list <- c(flip_list, df$pos_end[i])
      } else{
        in_progress <- FALSE
      }
    } else{
      in_progress <- FALSE
    }
  }
  return(flip_list)
}
```



Let's first take a look at a single pseudo-diploid and see what things look like.

```{r}
eagle_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/19april22_switch_errors/switch_errors/eagle/annotated"
shapeit_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/19april22_switch_errors/switch_errors/shapeit/annotated"
beagle_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/19april22_switch_errors/switch_errors/beagle/annotated"

id <- 10

s_file <- paste0(shapeit_dir, "/switch_meth_", id, ".csv")
e_file <- paste0(eagle_dir, "/switch_meth_", id, ".csv")
b_file <- paste0(beagle_dir, "/switch_meth_", id, ".csv")

s_df <- read_csv(s_file)
e_df <- read_csv(e_file)
b_df <- read_csv(b_file)

e_flip_list <- get_flip_pos(e_df)
s_flip_list <- get_flip_pos(s_df)
b_flip_list <- get_flip_pos(b_df)

s_df$is_flip <- as.numeric(s_df$pos_start %in% s_flip_list)
e_df$is_flip <- as.numeric(e_df$pos_start %in% e_flip_list)
b_df$is_flip <- as.numeric(b_df$pos_start %in% b_flip_list)
```

```{r}
s_df %>%
  filter(cpg_start == 1) %>%
  mutate(is_meth = meth > 0.5) %>%
  ggplot(aes(x = meth)) + geom_boxplot() +
  xlab("Methylation") +
  ggtitle("Distribution of Methylation Measures: SHAPEIT")

e_df %>%
  filter(cpg_start == 1) %>%
  mutate(is_meth = meth > 0.5) %>%
  ggplot(aes(x = meth)) + geom_boxplot() +
  xlab("Methylation") +
  ggtitle("Distribution of Methylation Measures: EAGLE")

b_df %>%
  filter(cpg_start == 1) %>%
  mutate(is_meth = meth > 0.5) %>%
  ggplot(aes(x = meth)) + geom_boxplot() +
  xlab("Methylation") + 
  ggtitle("Distribution of Methylation Measures: BEAGLE")


s_df %>%
  filter(cpg_start == 1) %>%
  mutate(is_meth = meth > 0.5) %>%
  pull(is_meth) %>% table(useNA = "ifany")

e_df %>%
  filter(cpg_start == 1) %>%
  mutate(is_meth = meth > 0.5) %>%
  pull(is_meth) %>% table(useNA = "ifany")

b_df %>%
  filter(cpg_start == 1) %>%
  mutate(is_meth = meth > 0.5) %>%
  pull(is_meth) %>% table(useNA = "ifany")

```

```{r}
meth_sum <- function(id, beagle_dir, eagle_dir, shapeit_dir, background_meth_rate){
  s_file <- paste0(shapeit_dir, "/switch_meth_", id, ".csv")
  e_file <- paste0(eagle_dir, "/switch_meth_", id, ".csv")
  b_file <- paste0(beagle_dir, "/switch_meth_", id, ".csv")
  
  s_df <- read_csv(s_file, show_col_types = FALSE)
  e_df <- read_csv(e_file, show_col_types = FALSE)
  b_df <- read_csv(b_file, show_col_types = FALSE)
  
  e_flip_list <- get_flip_pos(e_df)
  s_flip_list <- get_flip_pos(s_df)
  b_flip_list <- get_flip_pos(b_df)
  
  s_df$is_flip <- as.numeric(s_df$pos_start %in% s_flip_list)
  e_df$is_flip <- as.numeric(e_df$pos_start %in% e_flip_list)
  b_df$is_flip <- as.numeric(b_df$pos_start %in% b_flip_list)
  
  p_meth_s <- {s_df %>% filter(is_flip == 1, cpg_start == 1) %>% select(meth) %>% drop_na() %>% dim() %>% .[1]} / 
    {s_df %>% filter(is_flip == 1, cpg_start == 1) %>% dim() %>% .[1]}
  p_meth_e <- {e_df %>% filter(is_flip == 1, cpg_start == 1) %>% select(meth) %>% drop_na() %>% dim() %>% .[1]} / 
    {e_df %>% filter(is_flip == 1, cpg_start == 1) %>% dim() %>% .[1]}
  p_meth_b <- {b_df %>% filter(is_flip == 1, cpg_start == 1) %>% select(meth) %>% drop_na() %>% dim() %>% .[1]} / 
    {b_df %>% filter(is_flip == 1, cpg_start == 1) %>% dim() %>% .[1]}
  
  p_cpg_m_s <- {s_df %>% filter(cpg_start == 1) %>% select(meth) %>% drop_na() %>% dim() %>% .[1]} / 
    {s_df %>% filter(cpg_start == 1) %>% dim() %>% .[1]}
  p_cpg_m_e <- {e_df %>% filter(cpg_start == 1) %>% select(meth) %>% drop_na() %>% dim() %>% .[1]} / 
    {e_df %>% filter(cpg_start == 1) %>% dim() %>% .[1]}
  p_cpg_m_b <- {b_df %>% filter(cpg_start == 1) %>% select(meth) %>% drop_na() %>% dim() %>% .[1]} / 
    {b_df %>% filter(cpg_start == 1) %>% dim() %>% .[1]}
  
  min_meth_s <- s_df %>% pull(meth) %>% min(na.rm = T)
  min_meth_e <- e_df %>% pull(meth) %>% min(na.rm = T)
  min_meth_b <- b_df %>% pull(meth) %>% min(na.rm = T)
  
  p_is_meth_s <- s_df %>% select(meth) %>% drop_na() %>% {sum(.$meth > 0.5) / dim(.)[1]}
  p_is_meth_e <- e_df %>% select(meth) %>% drop_na() %>% {sum(.$meth > 0.5) / dim(.)[1]}
  p_is_meth_b <- b_df %>% select(meth) %>% drop_na() %>% {sum(.$meth > 0.5) / dim(.)[1]}
  
  e_meth_s <- p_is_meth_s / background_meth_rate
  e_meth_e <- p_is_meth_e / background_meth_rate
  e_meth_b <- p_is_meth_b / background_meth_rate
  
  p_flip_is_meth_s <- s_df %>% filter(cpg_start == 1) %>% select(meth) %>% drop_na() %>% {sum(.$meth > 0.5) / dim(.)[1]}
  p_flip_is_meth_e <- e_df %>% filter(cpg_start == 1) %>% select(meth) %>% drop_na() %>% {sum(.$meth > 0.5) / dim(.)[1]}
  p_flip_is_meth_b <- b_df %>% filter(cpg_start == 1) %>% select(meth) %>% drop_na() %>% {sum(.$meth > 0.5) / dim(.)[1]}
  
  return(data.frame(p_meth_s = p_meth_s,
                    p_meth_b = p_meth_b,
                    p_meth_e = p_meth_e,
                    min_meth_s = min_meth_s,
                    min_meth_e = min_meth_e,
                    min_meth_b = min_meth_b,
                    p_is_meth_s = p_is_meth_s,
                    p_is_meth_e = p_is_meth_e,
                    p_is_meth_b = p_is_meth_b,
                    p_cpg_m_s = p_cpg_m_s,
                    p_cpg_m_e = p_cpg_m_e,
                    p_cpg_m_b = p_cpg_m_b,
                    p_flip_is_meth_s = p_flip_is_meth_s,
                    p_flip_is_meth_e = p_flip_is_meth_e,
                    p_flip_is_meth_b = p_flip_is_meth_b,
                    e_meth_s = e_meth_s,
                    e_meth_e = e_meth_e,
                    e_meth_b = e_meth_b))
}
```

```{r}
meth_df <- lapply(c(1:300), 
                    function(x){
                      meth_sum(x, beagle_dir, eagle_dir, shapeit_dir, background_meth_rate)
                      }) %>%
  bind_rows()

meth_df$pop <- c(rep("EUR", 100), rep("AFR", 200))
```

```{r}
meth_df %>%
  select(pop, p_is_meth_s:p_is_meth_b) %>%
  rename(SHAPEIT = p_is_meth_s, BEAGLE = p_is_meth_b, EAGLE = p_is_meth_e) %>%
  pivot_longer(SHAPEIT:BEAGLE, names_to = "Method", values_to = "pct_is_meth") %>%
  ggplot(aes(x = pct_is_meth, colour = Method)) +
  geom_density(show.legend = FALSE) +
  stat_density(aes(x = pct_is_meth, colour = Method), geom="line", position = "identity", size = 0) +
  ggtitle("Percent Sites with Values > 0.5") +
  scale_colour_manual(values=cbPalette) +
  guides(colour = guide_legend(override.aes=list(size=1)))
```

```{r}
meth_df %>%
  select(pop, p_meth_s:p_meth_e) %>%
  rename(SHAPEIT = p_meth_s, BEAGLE = p_meth_b, EAGLE = p_meth_e) %>%
  pivot_longer(SHAPEIT:EAGLE, names_to = "Method", values_to = "pct_meth") %>%
  ggplot(aes(x = pct_meth, colour = Method)) +
  geom_density(show.legend = FALSE) +
  stat_density(aes(x = pct_meth, colour = Method), geom="line", position = "identity", size = 0) +
  ggtitle("Percent Sites with Values") +
  scale_colour_manual(values=cbPalette) +
  guides(colour = guide_legend(override.aes=list(size=1)))
```

