---
title: "NYGC_phasing_results_exploration"
author: "Andy Beck"
date: "2023-03-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(tidyverse)

base_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/nygc/"
switch_dir <- paste0(base_dir, "switch_errors/") # pattern: algorithm/annotated/switch_id.csv
het_dir <- paste0(base_dir, "het_loc/annotated/") # pattern: pair_id.csv
whats_dir <- paste0(base_dir, "whatshap/") # pattern: algorithm/eval_id.tsv
```


## Individual Pseudo-diploid Example

Let us first look at the results for a single pseudo-diploid. In particular, we will first look at the `vcftools` output, which lists the locations of differences between the phased VCF and the true-phase VCF. We'll utilize the following function to distinguish between switch and flip errors:

```{r}
get_flip_pos <- function(df){
  flip_list <- c()
  in_progress <- FALSE
  for(i in 1:(length(df$pos_start)-1)){
    if(df$pos_end[i] == df$pos_start[i+1]){
      if(!in_progress){
        in_progress <- TRUE
        flip_list <- c(flip_list, df$pos_end[i])
      } else{
        in_progress <- FALSE
      }
    } else{
      in_progress <- FALSE
    }
  }
  return(flip_list)
}
```

Let's first grab the locations of the EAGLE errors and the locations of heterozygous positions for the first pseudo-diploid:

```{r}
df_switch <- read_csv(paste0(switch_dir, "eagle", "/annotated/switch_", 1, ".csv"))
df_het <- read_csv(paste0(het_dir, "pair_", 1, ".csv"))

# Identify the flips among the errors:
flip_loc <- get_flip_pos(df_switch)
df_switch$flip_start <- df_switch$pos_start %in% flip_loc
df_switch$flip_end <- df_switch$pos_end %in% flip_loc
df_switch <- df_switch %>%
  mutate(flip = flip_end | flip_start)
```

And to verify our counts, let's grab the. results from whatshap for this pseudo-diploid:

```{r}
df_wh <- read_tsv(paste0(whats_dir, "eagle", "/eval_", 1, ".tsv"), show_col_types = FALSE)

df_wh$all_switchflips

sum(df_switch$flip != TRUE)
sum(df_switch$flip) / 2
```

Alrighty, now let's compare the proportion of flips and non-flip switches at CpG sites to the proportion of all het sites for this pd that are CpG:

```{r}
# proportion of total het sites at CpG:
sum(df_het$cpg)
length(df_het$cpg)
8535 / 59423 # around 14% of heterozygous sites in this PD are at CpG

sum(df_switch$cpg_start & df_switch$flip_end)
sum(df_switch$flip_end)
8 / 80 # eight of the eighty flips started at CpG sites

sum(df_switch$cpg_start & df_switch$flip != TRUE)
sum(df_switch$flip != TRUE)
18 / 112 # ~16% of non-flip switches are at CpG sites
```

Now let's write a function to get these stats for a single pseudo-diploid:

```{r}
get_cpg_switch_count <- function(alg, id, base_dir){
  switch_dir <- paste0(base_dir, "switch_errors/") # pattern: algorithm/annotated/switch_id.csv
  het_dir <- paste0(base_dir, "het_loc/annotated/") # pattern: pair_id.csv
  whats_dir <- paste0(base_dir, "whatshap/") # pattern: algorithm/eval_id.tsv
  
  df_switch <- read_csv(paste0(switch_dir, alg, "/annotated/switch_", id, ".csv"), show_col_types = FALSE)
  df_het <- read_csv(paste0(het_dir, "pair_", id, ".csv"), show_col_types = FALSE)
  df_wh <- read_tsv(paste0(whats_dir, alg, "/eval_", id, ".tsv"), show_col_types = FALSE)
  flip_loc <- get_flip_pos(df_switch)
  df_switch$flip_start <- df_switch$pos_start %in% flip_loc
  df_switch$flip_end <- df_switch$pos_end %in% flip_loc
  df_switch <- df_switch %>%
    mutate(flip = flip_end | flip_start)
  
  n_flips <- sum(df_switch$flip) / 2
  n_nf_switches <- sum(df_switch$flip != TRUE)
  
  n_nf_cpg <- sum(df_switch$cpg_start & df_switch$flip != TRUE)
  n_flip_cpg <- sum(df_switch$cpg_start & df_switch$flip_end)
  
  n_het_cpg <- sum(df_het$cpg)
  n_het <- length(df_het$cpg)
  
  return(data.frame(alg = alg, 
                    id = id, 
                    n_het = n_het,
                    n_het_cpg = n_het_cpg,
                    n_flip = n_flips,
                    n_switch = n_nf_switches,
                    n_flip_cpg = n_flip_cpg,
                    n_switch_cpg = n_nf_cpg))
}
```

And generate a data frame with all samples:

```{r}
get_cpg_switch_all <- function(alg, base_dir, N = 400){
  df <- get_cpg_switch_count(alg, 1, base_dir)
  for(i in 2:N){
    df <- bind_rows(df, get_cpg_switch_count(alg, i, base_dir))
  }
  return(df)
}
```

### EAGLE

Let's look at the CpG enrichment for EAGLE:

```{r}
df_e <- get_cpg_switch_all("eagle", base_dir)
```

Using these columns, let's compute enrichment of (flips/non-flips) at CpGs relative to hets at CpGs:

```{r}
df_e <- df_e %>%
  mutate(cpg_het = n_het_cpg / n_het,
         cpg_flip = n_flip_cpg / n_flip,
         cpg_switch = n_switch_cpg / n_switch) %>%
  mutate(enrich_flip = cpg_flip / cpg_het,
         enrich_switch = cpg_switch / cpg_het)
```

And let's make a boxplot:

```{r}
df_e %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pivot_longer(-id, names_to = "type", values_to = "enrich") %>%
  ggplot(aes(x = type, y = enrich)) +
  geom_boxplot() +
  ggtitle("Flip/Switch Enrichment at CpG", "EAGLE") +
  xlab("Error Type") +
  ylab("Enrichment Relative to Het. Site CpG Proportion")
```

For EAGLE, it looks like there's more of an enrichment for non-flip switches than for flips (albeit the mean of the flip enrichment is statistically significantly different from 1 according to a simple t test). Let's see if this pattern is consistent across the three phasing methods:

```{r}
df_b <- get_cpg_switch_all("beagle", base_dir)
df_s <- get_cpg_switch_all("shapeit", base_dir)


df_b <- df_b %>%
  mutate(cpg_het = n_het_cpg / n_het,
         cpg_flip = n_flip_cpg / n_flip,
         cpg_switch = n_switch_cpg / n_switch) %>%
  mutate(enrich_flip = cpg_flip / cpg_het,
         enrich_switch = cpg_switch / cpg_het)

df_s <- df_s %>%
  mutate(cpg_het = n_het_cpg / n_het,
         cpg_flip = n_flip_cpg / n_flip,
         cpg_switch = n_switch_cpg / n_switch) %>%
  mutate(enrich_flip = cpg_flip / cpg_het,
         enrich_switch = cpg_switch / cpg_het)

df_e$pop <- c(rep("EUR", 200), rep("AFR", 200))
df_b$pop <- c(rep("EUR", 200), rep("AFR", 200))
df_s$pop <- c(rep("EUR", 200), rep("AFR", 200))

```


```{r}
df_b %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pivot_longer(-id, names_to = "type", values_to = "enrich") %>%
  ggplot(aes(x = type, y = enrich)) +
  geom_boxplot() +
  ggtitle("Flip/Switch Enrichment at CpG", "BEAGLE") +
  xlab("Error Type") +
  ylab("Enrichment Relative to Het. Site CpG Proportion")
```

```{r}
df_s %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pivot_longer(-id, names_to = "type", values_to = "enrich") %>%
  ggplot(aes(x = type, y = enrich)) +
  geom_boxplot() +
  ggtitle("Flip/Switch Enrichment at CpG", "SHAPEIT") +
  xlab("Error Type") +
  ylab("Enrichment Relative to Het. Site CpG Proportion")
```

### Summary Statistics

```{r}
df_s %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pull(switches) %>%
  summary()

df_s %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pull(flips) %>%
  summary()

df_b %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pull(switches) %>%
  summary()

df_b %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pull(flips) %>%
  summary()

df_e %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pull(switches) %>%
  summary()

df_e %>% 
  select(id, starts_with("enrich")) %>%
  rename(flips = enrich_flip, switches = enrich_switch) %>%
  pull(flips) %>%
  summary()
```

## Comparisons Across Methods

### EAGLE and SHAPEIT

```{r}
df_s %>%
  select(id, n_flip, pop) %>%
  rename(shapeit = n_flip) %>%
  full_join({df_e %>% select(id, n_flip, pop) %>% rename(eagle = n_flip)}, by = c("id", "pop")) %>%
  ggplot(aes(x = shapeit, y = eagle, colour = pop)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  ggtitle("Flips by Method") + 
  theme_classic() +
  labs(color = "Population")

df_s %>%
  select(id, n_switch, pop) %>%
  rename(shapeit = n_switch) %>%
  full_join({df_e %>% select(id, n_switch, pop) %>% rename(eagle = n_switch)}, by = c("id", "pop")) %>%
  ggplot(aes(x = shapeit, y = eagle, colour = pop)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  ggtitle("Switches by Method") + 
  theme_classic() +
  labs(color = "Population") +
  guides(color = "none")
```

### BEAGLE and SHAPEIT

```{r}
df_s %>%
  select(id, n_flip, pop) %>%
  rename(shapeit = n_flip) %>%
  full_join({df_b %>% select(id, n_flip, pop) %>% rename(beagle = n_flip)}, by = c("id", "pop")) %>%
  ggplot(aes(x = shapeit, y = beagle, colour = pop)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  ggtitle("Flips by Method") + 
  theme_classic() +
  labs(color = "Population")

df_s %>%
  select(id, n_switch, pop) %>%
  rename(shapeit = n_switch) %>%
  full_join({df_b %>% select(id, n_switch, pop) %>% rename(beagle = n_switch)}, by = c("id", "pop")) %>%
  ggplot(aes(x = shapeit, y = beagle, colour = pop)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  ggtitle("Switches by Method") + 
  theme_classic() +
  labs(color = "Population") +
  guides(color = "none")
```

### BEAGLE and EAGLE

```{r}
df_e %>%
  select(id, n_flip, pop) %>%
  rename(eagle = n_flip) %>%
  full_join({df_b %>% select(id, n_flip, pop) %>% rename(beagle = n_flip)}, by = c("id", "pop")) %>%
  ggplot(aes(x = eagle, y = beagle, colour = pop)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  ggtitle("Flips by Method") + 
  theme_classic() +
  labs(color = "Population")

df_e %>%
  select(id, n_switch, pop) %>%
  rename(eagle = n_switch) %>%
  full_join({df_b %>% select(id, n_switch, pop) %>% rename(beagle = n_switch)}, by = c("id", "pop")) %>%
  ggplot(aes(x = eagle, y = beagle, colour = pop)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  ggtitle("Switches by Method") + 
  theme_classic() +
  labs(color = "Population") +
  guides(color = "none")
```


## Overlap of Errors

A natural question we may ask is in regards to the locations of the errors, and whether or not there are common locations between the algorithms. Let's first take a look at a single pseudo-diploid and see if there are any interesting metrics we'd like to pluck from all 400 of them.

```{r}
df_e <- read_csv(paste0(switch_dir, "eagle", "/annotated/switch_", 1, ".csv"), show_col_types = FALSE)
df_b <- read_csv(paste0(switch_dir, "beagle", "/annotated/switch_", 1, ".csv"), show_col_types = FALSE)
df_s <- read_csv(paste0(switch_dir, "shapeit", "/annotated/switch_", 1, ".csv"), show_col_types = FALSE)

# annotate with flips
flip_loc_e <- get_flip_pos(df_e)
df_e$flip_start <- df_e$pos_start %in% flip_loc_e
df_e$flip_end <- df_e$pos_end %in% flip_loc_e
df_e <- df_e %>%
  mutate(flip = flip_end | flip_start)

flip_loc_b <- get_flip_pos(df_b)
df_b$flip_start <- df_b$pos_start %in% flip_loc_b
df_b$flip_end <- df_b$pos_end %in% flip_loc_b
df_b <- df_b %>%
  mutate(flip = flip_end | flip_start)

flip_loc_s <- get_flip_pos(df_s)
df_s$flip_start <- df_s$pos_start %in% flip_loc_s
df_s$flip_end <- df_s$pos_end %in% flip_loc_s
df_s <- df_s %>%
  mutate(flip = flip_end | flip_start)
```

Okie dokey, let's see what the venn diagram of overlapping error locations looks like for this sample:

```{r}
library(ggVennDiagram)
x <- list(eagle = {df_e %>% filter(flip == FALSE) %>% pull(pos_start)},
          beagle = {df_b %>% filter(flip == FALSE) %>% pull(pos_start)},
          shapeit = {df_s %>% filter(flip == FALSE) %>% pull(pos_start)})
ggVennDiagram(x) +
  ggtitle("Error Location Overlap", "Non-flip Switches") +
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF") +
  theme(legend.position = c(.9, .9))
```

# Summarize Flip Locations over All Samples

```{r}
# df_e_final <- df_e %>%
#   select(pos_start, cpg_start, flip) %>%
#   mutate(switches = as.numeric(flip == FALSE),
#          flips = 1 - switches) %>%
#   select(pos_start, cpg_start, switches, flips)
# df_b_final <- df_b %>%
#   select(pos_start, cpg_start, flip) %>%
#   mutate(switches = as.numeric(flip == FALSE),
#          flips = 1 - switches) %>%
#   select(pos_start, cpg_start, switches, flips)
# df_s_final <- df_s %>%
#   select(pos_start, cpg_start, flip) %>%
#   mutate(switches = as.numeric(flip == FALSE),
#          flips = 1 - switches) %>%
#   select(pos_start, cpg_start, switches, flips)
# 
# for(i in 2:400){
#   df_e <- read_csv(paste0(switch_dir, "eagle", "/annotated/switch_", i, ".csv"), show_col_types = FALSE)
#   df_b <- read_csv(paste0(switch_dir, "beagle", "/annotated/switch_", i, ".csv"), show_col_types = FALSE)
#   df_s <- read_csv(paste0(switch_dir, "shapeit", "/annotated/switch_", i, ".csv"), show_col_types = FALSE)
#   
#   # annotate with flips
#   flip_loc_e <- get_flip_pos(df_e)
#   df_e$flip_start <- df_e$pos_start %in% flip_loc_e
#   df_e$flip_end <- df_e$pos_end %in% flip_loc_e
#   df_e <- df_e %>%
#     mutate(flip = flip_end | flip_start)
#   
#   flip_loc_b <- get_flip_pos(df_b)
#   df_b$flip_start <- df_b$pos_start %in% flip_loc_b
#   df_b$flip_end <- df_b$pos_end %in% flip_loc_b
#   df_b <- df_b %>%
#     mutate(flip = flip_end | flip_start)
#   
#   flip_loc_s <- get_flip_pos(df_s)
#   df_s$flip_start <- df_s$pos_start %in% flip_loc_s
#   df_s$flip_end <- df_s$pos_end %in% flip_loc_s
#   df_s <- df_s %>%
#     mutate(flip = flip_end | flip_start)
#   
#   df_e_final <- df_e %>%
#     select(pos_start, cpg_start, flip) %>%
#     mutate(switches = as.numeric(flip == FALSE),
#            flips = 1 - switches) %>%
#     select(pos_start, cpg_start, switches, flips) %>%
#     bind_rows(df_e_final) %>%
#     group_by(pos_start, cpg_start) %>%
#     summarize(switches = sum(switches),
#               flips = sum(flips))
#   
#   df_b_final <- df_b %>%
#     select(pos_start, cpg_start, flip) %>%
#     mutate(switches = as.numeric(flip == FALSE),
#            flips = 1 - switches) %>%
#     select(pos_start, cpg_start, switches, flips) %>%
#     bind_rows(df_b_final) %>%
#     group_by(pos_start, cpg_start) %>%
#     summarize(switches = sum(switches),
#               flips = sum(flips))
#   
#   df_s_final <- df_s %>%
#     select(pos_start, cpg_start, flip) %>%
#     mutate(switches = as.numeric(flip == FALSE),
#            flips = 1 - switches) %>%
#     select(pos_start, cpg_start, switches, flips) %>%
#     bind_rows(df_s_final) %>%
#     group_by(pos_start, cpg_start) %>%
#     summarize(switches = sum(switches),
#               flips = sum(flips))
# }
# 
# # save so as to not have to run above again
# write_csv(df_e_final, "/net/snowwhite/home/beckandy/research/phasing/output/nygc/summaries/all_e_pos.csv")
# write_csv(df_b_final, "/net/snowwhite/home/beckandy/research/phasing/output/nygc/summaries/all_b_pos.csv")
# write_csv(df_s_final, "/net/snowwhite/home/beckandy/research/phasing/output/nygc/summaries/all_s_pos.csv")

```

## Venn diagrams

```{r}

df_e_final <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/nygc/summaries/all_e_pos.csv")
df_s_final <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/nygc/summaries/all_s_pos.csv")
df_b_final <- read_csv("/net/snowwhite/home/beckandy/research/phasing/output/nygc/summaries/all_b_pos.csv")

x <- list(eagle = {df_e_final %>% pull(pos_start)},
          beagle = {df_b_final %>% pull(pos_start)},
          shapeit = {df_s_final %>% pull(pos_start)})
ggVennDiagram(x) +
  ggtitle("Error Location Overlap", "All Errors") +
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF") +
  theme(legend.position = c(.9, .9))
```

```{r}
x <- list(eagle = {df_e_final %>% filter(flips > 0) %>% pull(pos_start)},
          beagle = {df_b_final %>% filter(flips > 0) %>% pull(pos_start)},
          shapeit = {df_s_final %>% filter(flips > 0) %>% pull(pos_start)})
ggVennDiagram(x) +
  ggtitle("Error Location Overlap", "Flips") +
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF") +
  theme(legend.position = c(.9, .9))
```

```{r}
x <- list(eagle = {df_e_final %>% filter(switches > 0) %>% pull(pos_start)},
          beagle = {df_b_final %>% filter(switches > 0) %>% pull(pos_start)},
          shapeit = {df_s_final %>% filter(switches > 0) %>% pull(pos_start)})
ggVennDiagram(x) +
  ggtitle("Error Location Overlap", "Switches") +
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF") +
  theme(legend.position = c(.9, .9))
```


### EAGLE position histograms and figures

```{r}
df_e_final %>%
  ggplot(aes(x = flips, y = switches)) +
  geom_point(alpha = 1/10) +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Number of flips and switches at shared positions", "EAGLE") +
  theme_classic()

df_e_final %>%
  select(flips, switches) %>%
  group_by(flips, switches) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = flips, y = switches, fill = log10(n))) +
  geom_tile() +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  theme_classic()
```

```{r}
df_e_final %>%
  filter(flips > 0) %>%
  ggplot(aes(x = flips)) +
  geom_histogram(bins = length(unique(df_e_final$flips))) +
  ggtitle("Number of flips at positions", "EAGLE") +
  xlab("Number of Pseudo-diploids") +
  ylab("Positions") +
  theme_classic()
```

## Bins

### Replication Timing Data

This data comes from https://www.thekorenlab.org/data

```{r}
df_rep <- read_tsv("data/rep_timing.txt") %>%
  rename(rt = `Replication Timing`) %>%
  mutate(bin = ceiling(Coordinate / 1000)) %>%
  filter(Chr == 23)
```

Average RT value per 1Mb bin:

```{r}
df_rep %>%
  group_by(bin) %>%
  summarize(mean_rt = mean(rt),
            min_rt = min(rt),
            max_rt = max(rt)) %>%
  ggplot(aes(x = bin,
             y = mean_rt)) +
  geom_line()
```

### Binning the Errors

#### EAGLE

```{r}
df_e_final %>%
  select(pos_start) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  select(bin) %>%
  group_by(bin) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = bin, y = n)) +
  geom_point() +
  theme_classic() +
  ggtitle("Positions with Errors per Bin", "EAGLE")
```

```{r}
df_e_final %>%
  select(pos_start, switches) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(switches)) %>%
  ggplot(aes(x = bin, y = n)) +
  geom_point() +
  theme_classic() +
  ggtitle("Switches per Bin", "EAGLE")
```

```{r}
df_e_final %>%
  select(pos_start, flips) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(flips)) %>%
  ggplot(aes(x = bin, y = n)) +
  geom_point() +
  theme_classic() +
  ggtitle("Flips per Bin", "EAGLE")
```

```{r}
coeff <- 30.25
df_e_final %>%
  select(pos_start, switches) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(switches)) %>%
  full_join(df_rep, by = "bin") %>%
  ggplot(aes(x = bin)) +
  geom_point(aes(y = n)) +
  geom_line(aes(y = (rt + 2) * coeff), color = "red") +
  ylim(c(-2,2)) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Number of Switches",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(./coeff)-2, name="Replication Timing")
  ) + 
  
  theme_classic() +
  ggtitle("Switches per MB Bin", "EAGLE")
```

```{r}
coeff <- 25.25
df_e_final %>%
  select(pos_start, flips) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(flips)) %>%
  full_join(df_rep, by = "bin") %>%
  ggplot(aes(x = bin)) +
  geom_point(aes(y = n)) +
  geom_line(aes(y = (rt + 2) * coeff), color = "red") +
  ylim(c(-2,2)) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Number of Flips",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(./coeff)-2, name="Replication Timing")
  ) + 
  
  theme_classic() +
  ggtitle("Flips per MB Bin", "EAGLE")
```

```{r}
coeff <- 10
df_e_final %>%
  select(pos_start) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  select(bin) %>%
  group_by(bin) %>%
  summarise(n = n()) %>%
  full_join(df_rep, by = "bin") %>%
  ggplot(aes(x = bin)) +
  geom_line(aes(y = rt), color = "red") +
  geom_line(aes(y = (n-10) / coeff)) +
  ylim(c(-2,2)) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Replication Timing",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(.*coeff)+10, name="Number of Errors")
  ) + 
  
  theme_classic() +
  ggtitle("Positions with Errors per MB Bin", "EAGLE")
```

#### BEAGLE

```{r}
df_b_final %>%
  select(pos_start, switches) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(switches)) %>%
  ggplot(aes(x = bin, y = n)) +
  geom_point() +
  theme_classic() +
  ggtitle("Switches per Bin", "BEAGLE")
```

```{r}
coeff <- 33.25
df_b_final %>%
  select(pos_start, switches) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(switches)) %>%
  full_join(df_rep, by = "bin") %>%
  ggplot(aes(x = bin)) +
  geom_point(aes(y = n)) +
  geom_line(aes(y = (rt + 2) * coeff), color = "red") +
  ylim(c(-2,2)) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Number of Switches",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(./coeff)-2, name="Replication Timing")
  ) + 
  
  theme_classic() +
  ggtitle("Switches per MB Bin", "BEAGLE")
```

```{r}
coeff <- 24.75
df_b_final %>%
  select(pos_start, flips) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(flips)) %>%
  full_join(df_rep, by = "bin") %>%
  ggplot(aes(x = bin)) +
  geom_point(aes(y = n)) +
  geom_line(aes(y = (rt + 2) * coeff), color = "red") +
  ylim(c(-2,2)) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Number of Flips",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(./coeff)-2, name="Replication Timing")
  ) + 
  
  theme_classic() +
  ggtitle("Flips per MB Bin", "BEAGLE")
```

#### SHAPEIT

```{r}
coeff <- 28.25
df_s_final %>%
  select(pos_start, switches) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(switches)) %>%
  full_join(df_rep, by = "bin") %>%
  ggplot(aes(x = bin)) +
  geom_point(aes(y = n)) +
  geom_line(aes(y = (rt + 2) * coeff), color = "red") +
  ylim(c(-2,2)) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Number of Switches",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(./coeff)-2, name="Replication Timing")
  ) + 
  
  theme_classic() +
  ggtitle("Switches per MB Bin", "SHAPEIT")
```

```{r}
coeff <- 26.5
df_s_final %>%
  select(pos_start, flips) %>%
  mutate(bin = ceiling(pos_start / 1000)) %>%
  group_by(bin) %>%
  summarise(n = sum(flips)) %>%
  full_join(df_rep, by = "bin") %>%
  ggplot(aes(x = bin)) +
  geom_point(aes(y = n)) +
  geom_line(aes(y = (rt + 2) * coeff), color = "red") +
  ylim(c(-2,2)) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Number of Flips",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(./coeff)-2, name="Replication Timing")
  ) + 
  
  theme_classic() +
  ggtitle("Flips per MB Bin", "SHAPEIT")
```

