---
title: "venn_diagram"
author: "Andy Beck"
date: "2024-02-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r, include=FALSE}
library(tidyverse)
library(ggExtra)
library(ggVennDiagram)
library(limma)
source("code/common_functions.R")
'%!in%' <- function(x,y)!('%in%'(x,y))
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
eagle_switch_dir <-   "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/switch_errors/eagle/annotated/"
shapeit_switch_dir <- "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/switch_errors/shapeit/annotated/"
beagle_switch_dir <-  "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/switch_errors/beagle/annotated/"
num_sites_dir <-      "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/vcf_n_sites/"
whatshap_dir <-       "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/whatshap/"
het_loc_dir <-        "/net/snowwhite/home/beckandy/research/phasing/output/final_switch_errors/het_loc/annotated/"
pair_info_df <- read_delim("data/sample_pairs_9aug2023.csv", col_names = c("POP", "ID1", "ID2"))
pair_info_df$SP <- c(rep("EUR", 200), rep("AFR", 200), rep("AMR", 100), rep("EAS", 100), rep("SAS", 100))
df_freq <- read_tsv("/net/snowwhite/home/beckandy/research/phasing/output/chrX_maf_simple.tsv")
```

## Introduction

In this document we will explore the construction of a "venn diagram" of the overlap of phasing errors across three methods over 700 synthetic diploids. Intuitively, this makes sense for a single synthetic diploid: we partition all errors into categories determined by which method(s) produced the error. The potential complication is if we wish to look only at particular classes of errors, e.g. switches or flips. For an example of what can happen, suppose a heterozygous location is the site of a switch in Beagle. That same position could be the location of a switch in Eagle or SHAPEIT, or it may have been correctly phased. Those are not exhaustive, however; another potential outcome is that same location is the latter half of a flip error in either Eagle or SHAPEIT. Is this still considered an overlapping switch?

For a first pass, I'll be a bit exhaustive in terms of the categories of overlap for the errors in a single synthetic diploid.

```{r}
load_het_pos <- function(id, het_loc_dir){
  fname <- paste0(het_loc_dir, "pair_", id, ".csv")
  return(read_csv(fname, show_col_types = FALSE))
}

load_switch_df <- function(id, switch_dir){
  fname <- paste0(switch_dir, "switch_", id, ".csv")
  results <- read_csv(fname, show_col_types = FALSE)
  flip_loc <- get_flip_pos(results)
  results$is_flip <- (results$pos_start %in% flip_loc) | (results$pos_end %in% flip_loc)
  results$flip <- results$pos_end %in% flip_loc
  # results <- results %>%
  #   filter(flip | (!is_flip & !flip))
  # return(results %>% select(-is_flip))
  return(results)
}

annotate_het_list <- function(id, het_loc_dir, beagle_dir, eagle_dir, shapeit_dir, df_freq){
  result <- load_het_pos(id, het_loc_dir)
  df_b <- load_switch_df(id, beagle_dir)
  df_e <- load_switch_df(id, eagle_dir)
  df_s <- load_switch_df(id, shapeit_dir)
  
  result$switch_b <- result$pos %in% {df_b %>% filter(!is_flip) %>% pull(pos_start)}
  result$flip_b <- result$pos %in% {df_b %>% filter(flip) %>% pull(pos_start)}
  result$flip2_b <- result$pos %in% {df_b %>% filter(is_flip & !flip) %>% pull(pos_start)}
  
  result$switch_e <- result$pos %in% {df_e %>% filter(!is_flip) %>% pull(pos_start)}
  result$flip_e <- result$pos %in% {df_e %>% filter(flip) %>% pull(pos_start)}
  result$flip2_e <- result$pos %in% {df_e %>% filter(is_flip & !flip) %>% pull(pos_start)}
  
  result$switch_s <- result$pos %in% {df_s %>% filter(!is_flip) %>% pull(pos_start)}
  result$flip_s <- result$pos %in% {df_s %>% filter(flip) %>% pull(pos_start)}
  result$flip2_s <- result$pos %in% {df_s %>% filter(is_flip & !flip) %>% pull(pos_start)}
  
  result <- left_join(result,
                       {df_freq %>%
                           select(POS, maf)},
                       by = c("pos" = "POS"))
  
  return(result)
}

df <- annotate_het_list(1, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)
```

For the `ggVennDiagram` library, we will want lists of positions for each error type (note that here we "double count" flips by including the positions of both involved heterozygous sites):

```{r}
x <- list(beagle = {df %>% filter(switch_b | flip_b | flip2_b) %>% pull(pos)},
          eagle = {df %>% filter(switch_e | flip_e | flip2_e) %>% pull(pos)},
          shapeit = {df %>% filter(switch_s | flip_s | flip2_s) %>% pull(pos)})
ggVennDiagram(x, label = "count") + scale_fill_gradient(low="grey90",high = "red")
```

I want to be a bit more rigorous in counting overlaps; for each method, I only want start sites to overlap errors of other methods, i.e. I don't want to double count flips in method A, but would count as overlap if same position were the second-half of a flip in method B.

```{r}
# Get flip and switch positions
b_flip_pos <- {df %>% filter(flip_b) %>% pull(pos)}
e_flip_pos <- {df %>% filter(flip_e) %>% pull(pos)}
s_flip_pos <- {df %>% filter(flip_s) %>% pull(pos)}
b_switch_pos <- {df %>% filter(switch_b) %>% pull(pos)}
e_switch_pos <- {df %>% filter(switch_e) %>% pull(pos)}
s_switch_pos <- {df %>% filter(switch_s) %>% pull(pos)}
b_flip2_pos <- {df %>% filter(flip2_b) %>% pull(pos)}
e_flip2_pos <- {df %>% filter(flip2_e) %>% pull(pos)}
s_flip2_pos <- {df %>% filter(flip2_s) %>% pull(pos)}

# Beagle overlap
b_switch_unique <- b_switch_pos[b_switch_pos %!in% c(e_flip_pos, s_flip_pos, e_switch_pos, s_switch_pos, e_flip2_pos, s_flip2_pos)]
b_switch_e <- b_switch_pos[(b_switch_pos %in% c(e_flip_pos, e_switch_pos, e_flip2_pos)) & b_switch_pos %!in% c(s_flip_pos, s_switch_pos, s_flip2_pos)]
b_switch_s <- b_switch_pos[(b_switch_pos %in% c(s_flip_pos, s_switch_pos, s_flip2_pos)) & b_switch_pos %!in% c(e_flip_pos, e_switch_pos, e_flip2_pos)]
b_switch_e_s <- b_switch_pos[b_switch_pos %in% c(s_flip_pos, s_switch_pos, s_flip2_pos) & b_switch_pos %in% c(e_flip_pos, e_switch_pos, e_flip2_pos)]

b_flip_unique <- b_flip_pos[b_flip_pos %!in% c(e_flip_pos, s_flip_pos, e_switch_pos, s_switch_pos, e_flip2_pos, s_flip2_pos)]
b_flip_e <- b_flip_pos[(b_flip_pos %in% c(e_flip_pos, e_switch_pos, e_flip2_pos)) & b_flip_pos %!in% c(s_flip_pos, s_switch_pos, s_flip2_pos)]
b_flip_s <- b_flip_pos[(b_flip_pos %in% c(s_flip_pos, s_switch_pos, s_flip2_pos)) & b_flip_pos %!in% c(e_flip_pos, e_switch_pos, e_flip2_pos)]
b_flip_e_s <- b_flip_pos[b_flip_pos %in% c(s_flip_pos, s_switch_pos, s_flip2_pos) & b_flip_pos %in% c(e_flip_pos, e_switch_pos, e_flip2_pos)]

# Eagle Overlap
e_switch_unique <- e_switch_pos[e_switch_pos %!in% c(b_flip_pos, s_flip_pos, b_switch_pos, s_switch_pos, b_flip2_pos, s_flip2_pos)]
e_switch_b <- e_switch_pos[(e_switch_pos %in% c(b_flip_pos, b_switch_pos, b_flip2_pos)) & e_switch_pos %!in% c(s_flip_pos, s_switch_pos, s_flip2_pos)]
e_switch_s <- e_switch_pos[(e_switch_pos %in% c(s_flip_pos, s_switch_pos, s_flip2_pos)) & e_switch_pos %!in% c(b_flip_pos, b_switch_pos, b_flip2_pos)]
e_switch_b_s <- e_switch_pos[e_switch_pos %in% c(s_flip_pos, s_switch_pos, s_flip2_pos) & e_switch_pos %in% c(b_flip_pos, b_switch_pos, b_flip2_pos)]

e_flip_unique <- e_flip_pos[e_flip_pos %!in% c(b_flip_pos, s_flip_pos, b_switch_pos, s_switch_pos, b_flip2_pos, s_flip2_pos)]
e_flip_b <- e_flip_pos[(e_flip_pos %in% c(b_flip_pos, b_switch_pos, b_flip2_pos)) & e_flip_pos %!in% c(s_flip_pos, s_switch_pos, s_flip2_pos)]
e_flip_s <- e_flip_pos[(e_flip_pos %in% c(s_flip_pos, s_switch_pos, s_flip2_pos)) & e_flip_pos %!in% c(b_flip_pos, b_switch_pos, b_flip2_pos)]
e_flip_b_s <- e_flip_pos[e_flip_pos %in% c(s_flip_pos, s_switch_pos, s_flip2_pos) & e_flip_pos %in% c(b_flip_pos, b_switch_pos, b_flip2_pos)]

# SHAPEIT Overlap
s_switch_unique <- s_switch_pos[s_switch_pos %!in% c(b_flip_pos, e_flip_pos, b_switch_pos, e_switch_pos, b_flip2_pos, e_flip2_pos)]
s_switch_b <- s_switch_pos[(s_switch_pos %in% c(b_flip_pos, b_switch_pos, b_flip2_pos)) & s_switch_pos %!in% c(e_flip_pos, e_switch_pos, e_flip2_pos)]
s_switch_e <- s_switch_pos[(s_switch_pos %in% c(e_flip_pos, e_switch_pos, e_flip2_pos)) & s_switch_pos %!in% c(b_flip_pos, b_switch_pos, b_flip2_pos)]
s_switch_b_e <- s_switch_pos[s_switch_pos %in% c(e_flip_pos, e_switch_pos, e_flip2_pos) & s_switch_pos %in% c(b_flip_pos, b_switch_pos, b_flip2_pos)]

s_flip_unique <- s_flip_pos[s_flip_pos %!in% c(e_flip_pos, b_flip_pos, e_switch_pos, b_switch_pos, e_flip2_pos, b_flip2_pos)]
s_flip_b <- s_flip_pos[(s_flip_pos %in% c(b_flip_pos, b_switch_pos, b_flip2_pos)) & s_flip_pos %!in% c(e_flip_pos, e_switch_pos, e_flip2_pos)]
s_flip_e <- s_flip_pos[(s_flip_pos %in% c(e_flip_pos, e_switch_pos, e_flip2_pos)) & s_flip_pos %!in% c(b_flip_pos, b_switch_pos, b_flip2_pos)]
s_flip_b_e <- s_flip_pos[s_flip_pos %in% c(e_flip_pos, e_switch_pos, e_flip2_pos) & s_flip_pos %in% c(b_flip_pos, b_switch_pos, b_flip2_pos)]

# Get counts
```

I think I have to double count flips for this analysis, since we'd want to identify cases where the latter half of a flip is also the location of either a switch or the start of a flip in another method.

1. Switches: get list of which ones do or do not appear in the error lists for the other methods
2. Flips: identify overlaps, subsetting list to begin position

What's missing from above code is checking if flip2 overlaps an error in the other methods


Or, hear me out: we say nuts to the above, and just do the very simple criteria of:

1. Switches are shared only with switches that occur at the same position
2. Flips are shared only with flips that start at the same position
3. Everything else is distinct as far as we care

So for each SD we get two venn diagrams with 3 circles.

```{r}
x <- list(beagle = {df %>% filter(switch_b) %>% pull(pos)},
          eagle = {df %>% filter(switch_e) %>% pull(pos)},
          shapeit = {df %>% filter(switch_s) %>% pull(pos)})
p1 <- ggVennDiagram(x, label = "count") + 
  scale_fill_gradient(low="grey90",high = "green") + 
  ggtitle("Switches") + 
  theme(plot.title = element_text(hjust = 0.5))

x <- list(beagle = {df %>% filter(flip_b) %>% pull(pos)},
          eagle = {df %>% filter(flip_e) %>% pull(pos)},
          shapeit = {df %>% filter(flip_s) %>% pull(pos)})
p2 <- ggVennDiagram(x, label="count") + 
  scale_fill_gradient(low="grey90",high = "blue") + 
  ggtitle("Flips") + 
  theme(plot.title = element_text(hjust = 0.5))

gridExtra::grid.arrange(p1,p2,ncol=2)
```

Naturally, we may ask how many intersections of errors do we miss by considering only matching error types and, in the case of flips, only matching directly overlapping errors. First, we look at how many switches from each method overlap flips from the other two methods:

```{r}
x <- list(beagle = {df %>% filter(switch_b) %>% pull(pos)},
          eagle_flip = {df %>% filter(flip_e | flip2_e) %>% pull(pos)},
          eagle_switch = {df %>% filter(switch_e) %>% pull(pos)},
          shapeit_switch = {df %>% filter(switch_s) %>% pull(pos)},
          shapeit_flip = {df %>% filter(flip_s | flip2_s) %>% pull(pos)})
p1 <- ggVennDiagram(x, label = "count") + scale_fill_gradient(low="grey90",high = "purple") + 
  theme(plot.title = element_text(hjust = 0.5))

x <- list(beagle = {df %>% filter(switch_b) %>% pull(pos)},
          eagle_flip = {df %>% filter(flip_e | flip2_e) %>% pull(pos)},
          shapeit_flip = {df %>% filter(flip_s | flip2_s) %>% pull(pos)})
p2 <- ggVennDiagram(x, label = "count") + scale_fill_gradient(low="grey90",high = "darkgreen") + 
  theme(plot.title = element_text(hjust = 0.5))

gridExtra::grid.arrange(p1,p2,ncol=2)
```

```{r}
x <- list(beagle_flip = {df %>% filter(flip_b | flip2_b) %>% pull(pos)},
          beagle_switch = {df %>% filter(switch_b) %>% pull(pos)},
          eagle = {df %>% filter(switch_e) %>% pull(pos)},
          shapeit_switch = {df %>% filter(switch_s) %>% pull(pos)},
          shapeit_flip = {df %>% filter(flip_s | flip2_s) %>% pull(pos)})
p1 <- ggVennDiagram(x, label = "count") + scale_fill_gradient(low="grey90",high = "purple") + 
  theme(plot.title = element_text(hjust = 0.5))

x <- list(beagle_flip = {df %>% filter(flip_b | flip2_b) %>% pull(pos)},
          eagle = {df %>% filter(switch_e) %>% pull(pos)},
          shapeit_flip = {df %>% filter(flip_s | flip2_s) %>% pull(pos)})
p2 <- ggVennDiagram(x, label = "count") + scale_fill_gradient(low="grey90",high = "darkgreen") + 
  theme(plot.title = element_text(hjust = 0.5))

gridExtra::grid.arrange(p1,p2,ncol=2)
```

```{r}
x <- list(beagle_flip = {df %>% filter(flip_b | flip2_b) %>% pull(pos)},
          beagle_switch = {df %>% filter(switch_b) %>% pull(pos)},
          eagle_switch = {df %>% filter(switch_e) %>% pull(pos)},
          eagle_flip = {df %>% filter(flip_e | flip2_e) %>% pull(pos)},
          shapeit = {df %>% filter(switch_s) %>% pull(pos)})
p1 <- ggVennDiagram(x, label = "count") + scale_fill_gradient(low="grey90",high = "purple") + 
  theme(plot.title = element_text(hjust = 0.5))

x <- list(beagle_flip = {df %>% filter(flip_b | flip2_b) %>% pull(pos)},
          eagle_flip = {df %>% filter(flip_e | flip2_e) %>% pull(pos)},
          shapeit = {df %>% filter(switch_s) %>% pull(pos)})
p2 <- ggVennDiagram(x, label = "count") + scale_fill_gradient(low="grey90",high = "darkgreen") + 
  theme(plot.title = element_text(hjust = 0.5))

gridExtra::grid.arrange(p1,p2,ncol=2)
```


So, if we wanted to aggregate these counts across multiple SDs, we'd have to generate 14 numbers per SD. Is there a quick way of doing this?

```{r}
df_switch <- df %>%
  filter(switch_b | switch_e | switch_s) %>%
  mutate(beagle = ifelse(switch_b, "Y", "N"),
         eagle = ifelse(switch_e, "Y", "N"),
         shapeit = ifelse(switch_s, "Y", "N")) %>%
  mutate(count_cat = paste(beagle, eagle, shapeit, sep="_")) %>%
  pull(count_cat) %>%
  table() %>%
  as.data.frame() %>%
  rename(count_cat = ".")

df_flip <- df %>%
  filter(flip_b | flip_e | flip_s) %>%
  mutate(beagle = ifelse(flip_b, "Y", "N"),
         eagle = ifelse(flip_e, "Y", "N"),
         shapeit = ifelse(flip_s, "Y", "N")) %>%
  mutate(count_cat = paste(beagle, eagle, shapeit, sep="_")) %>%
  pull(count_cat) %>%
  table() %>%
  as.data.frame() %>%
  rename(count_cat = ".")
```



## Simple Overlap

```{r}
# Version which only does exact matches of flips and switches
get_error_counts <- function(ids, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq){
  n <- length(ids)
  results <- vector(mode = "list", length = 2)
  
  df <- annotate_het_list(ids[1], het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)
  df_switch <- df %>%
    filter(switch_b | switch_e | switch_s) %>%
    mutate(beagle = ifelse(switch_b, "Y", "N"),
           eagle = ifelse(switch_e, "Y", "N"),
           shapeit = ifelse(switch_s, "Y", "N")) %>%
    mutate(count_cat = paste(beagle, eagle, shapeit, sep="_")) %>%
    pull(count_cat) %>%
    table() %>%
    as.data.frame() %>%
    rename(count_cat = ".")
  
  df_flip <- df %>%
    filter(flip_b | flip_e | flip_s) %>%
    mutate(beagle = ifelse(flip_b, "Y", "N"),
           eagle = ifelse(flip_e, "Y", "N"),
           shapeit = ifelse(flip_s, "Y", "N")) %>%
    mutate(count_cat = paste(beagle, eagle, shapeit, sep="_")) %>%
    pull(count_cat) %>%
    table() %>%
    as.data.frame() %>%
    rename(count_cat = ".")
  
  for(i in 2:n){
    df <- annotate_het_list(ids[i], het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)
    df_switch2 <- df %>%
      filter(switch_b | switch_e | switch_s) %>%
      mutate(beagle = ifelse(switch_b, "Y", "N"),
             eagle = ifelse(switch_e, "Y", "N"),
             shapeit = ifelse(switch_s, "Y", "N")) %>%
      mutate(count_cat = paste(beagle, eagle, shapeit, sep="_")) %>%
      pull(count_cat) %>%
      table() %>%
      as.data.frame() %>%
      rename(count_cat = ".", Freq2 = Freq)
    
    df_switch <- df_switch %>%
      full_join(df_switch2, by="count_cat") %>%
      replace_na(list("Freq" = 0, "Freq2" = 0)) %>%
      mutate(Freq = Freq + Freq2) %>%
      select(-Freq2)
      
    
    df_flip2 <- df %>%
      filter(flip_b | flip_e | flip_s) %>%
      mutate(beagle = ifelse(flip_b, "Y", "N"),
             eagle = ifelse(flip_e, "Y", "N"),
             shapeit = ifelse(flip_s, "Y", "N")) %>%
      mutate(count_cat = paste(beagle, eagle, shapeit, sep="_")) %>%
      pull(count_cat) %>%
      table() %>%
      as.data.frame() %>%
      rename(count_cat = ".", Freq2 = Freq)
    
    df_flip <- df_flip %>%
      full_join(df_flip2, by="count_cat") %>%
      replace_na(list("Freq" = 0, "Freq2" = 0)) %>%
      mutate(Freq = Freq + Freq2) %>%
      select(-Freq2)
  }
  results[[1]] <- df_switch
  results[[2]] <- df_flip
  return(results)
}
```

### EUR

```{r}
results_eur <- get_error_counts(1:200, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

results_eur[[1]] %>%
  knitr::kable()
results_eur[[2]] %>%
  knitr::kable()
```

```{r}
count_to_ven <- function(x){
  df <- x %>%
    mutate(beagle = ifelse(str_sub(count_cat, 1, 1)=="Y", 1, 0),
           eagle = ifelse(str_sub(count_cat, 3, 3)=="Y", 1, 0),
           shapeit = ifelse(str_sub(count_cat, 5, 5)== "Y", 1, 0)) %>%
    select(beagle, eagle, shapeit, Freq)
  
  n_b <- df %>% filter(beagle == 1, eagle == 0, shapeit == 0) %>% pull(Freq)
  n_e <- df %>% filter(beagle == 0, eagle == 1, shapeit == 0) %>% pull(Freq)
  n_s <- df %>% filter(beagle == 0, eagle == 0, shapeit == 1) %>% pull(Freq)
  n_b_e <- df %>% filter(beagle == 1, eagle == 1, shapeit == 0) %>% pull(Freq)
  n_b_s <- df %>% filter(beagle == 1, eagle == 0, shapeit == 1) %>% pull(Freq)
  n_e_s <- df %>% filter(beagle == 0, eagle == 1, shapeit == 1) %>% pull(Freq)
  n_b_e_s <- df %>% filter(beagle == 1, eagle == 1, shapeit == 1) %>% pull(Freq)
  
  data.frame(beagle = c(rep(TRUE, n_b), rep(FALSE, n_e),rep(FALSE, n_s), rep(TRUE, n_b_e), rep(TRUE, n_b_s), rep(FALSE, n_e_s) , rep(TRUE, n_b_e_s)),
             eagle = c(rep(FALSE, n_b), rep(TRUE, n_e),rep(FALSE, n_s), rep(TRUE, n_b_e), rep(FALSE, n_b_s), rep(TRUE, n_e_s) , rep(TRUE, n_b_e_s)),
             shapeit = c(rep(FALSE, n_b), rep(FALSE, n_e), rep(TRUE, n_s), rep(FALSE, n_b_e), rep(TRUE, n_b_s), rep(TRUE, n_e_s) , rep(TRUE, n_b_e_s)) 
             ) %>%
    vennCounts() %>%
    vennDiagram() %>% print()
}

count_to_ven(results_eur[[1]])
count_to_ven(results_eur[[2]])
```


### AFR

```{r}
results_afr <- get_error_counts(201:400, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

results_afr[[1]] %>%
  knitr::kable()
results_afr[[2]] %>%
  knitr::kable()

count_to_ven(results_afr[[1]])
count_to_ven(results_afr[[2]])
```

### AMR

```{r}
results_amr <- get_error_counts(401:500, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

results_amr[[1]] %>%
  knitr::kable()
results_amr[[2]] %>%
  knitr::kable()

count_to_ven(results_amr[[1]])
count_to_ven(results_amr[[2]])
```

### EAS

```{r}
results_eas <- get_error_counts(501:600, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

results_eas[[1]] %>%
  knitr::kable()
results_eas[[2]] %>%
  knitr::kable()

count_to_ven(results_eas[[1]])
count_to_ven(results_eas[[2]])
```

### SAS

```{r}
results_sas <- get_error_counts(601:700, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

results_sas[[1]] %>%
  knitr::kable()
results_sas[[2]] %>%
  knitr::kable()

count_to_ven(results_sas[[1]])
count_to_ven(results_sas[[2]])
```

# Errors in Bins

To get the sense of how much the methods overlap in a larger scale, we'll look at the aggregate number of errors in bins across the genome for both switches and flips. First, let's take a look at what this overlap looks like within a single synthetic diploid:

```{r}
bin_size <- 1e6 # try 1MB
df <- annotate_het_list(1, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq) %>%
  mutate(bin_id = ceiling(pos / bin_size))
```

Let's first look at switches across bins

```{r}
df %>%
  group_by(bin_id) %>%
  summarize(beagle = sum(switch_b),
            eagle = sum(switch_e),
            shapeit = sum(switch_s)) %>%
  pivot_longer(-bin_id, names_to = "method", values_to = "switches") %>%
  ggplot(aes(x = bin_id, y = switches, color = method)) +
  geom_point() +
  geom_line() +
  theme_classic()
```

Now let's lower our resolution and look at 10MB bins:

```{r}
bin_size <- 1e7
df %>%
  mutate(bin_id = ceiling(pos / bin_size)) %>%
  group_by(bin_id) %>%
  summarize(beagle = sum(switch_b),
            eagle = sum(switch_e),
            shapeit = sum(switch_s)) %>%
  pivot_longer(-bin_id, names_to = "method", values_to = "switches") %>%
  ggplot(aes(x = bin_id, y = switches, color = method)) +
  geom_point() +
  geom_line() +
  theme_classic()
```

## EUR

Now let's see what this distribution looks like across all EUR SDs. First, on the MB scale:

### Switches

```{r}
get_errors_mb <- function(ids, bin_size, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq){
  
  final_df <- annotate_het_list(ids[1], het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq) %>%
    mutate(bin_id = ceiling(pos / bin_size)) %>%
    group_by(bin_id) %>%
    summarize(beagle_s = sum(switch_b),
              eagle_s = sum(switch_e),
              shapeit_s = sum(switch_s),
              beagle_f = sum(flip_b),
              eagle_f = sum(flip_e),
              shapeit_f = sum(flip_s),
              hets = n())
  
  for(i in 2:length(ids)){
    df2 <- annotate_het_list(ids[i], het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq) %>%
      mutate(bin_id = ceiling(pos / bin_size)) %>%
      group_by(bin_id) %>%
      summarize(beagle_s2 = sum(switch_b),
                eagle_s2 = sum(switch_e),
                shapeit_s2 = sum(switch_s),
                beagle_f2 = sum(flip_b),
              eagle_f2 = sum(flip_e),
              shapeit_f2 = sum(flip_s),
              hets2 = n())
    final_df <- full_join(final_df, df2, by="bin_id") %>%
      replace_na(list("beagle_s" = 0, "eagle_s" = 0, "shapeit_s" = 0,
                      "beagle_s2" = 0, "eagle_s2" = 0, "shapeit_s2" = 0,
                      "beagle_f" = 0, "eagle_f" = 0, "shapeit_f" = 0,
                      "beagle_f2" = 0, "eagle_f2" = 0, "shapeit_f2" = 0,
                      "hets" = 0, "hets2" = 0)) %>%
      mutate(beagle_s = beagle_s + beagle_s2,
             eagle_s = eagle_s + eagle_s2,
             shapeit_s = shapeit_s + shapeit_s2,
             beagle_f = beagle_f + beagle_f2,
             eagle_f = eagle_f + eagle_f2,
             shapeit_f = shapeit_f + shapeit_f2,
             hets = hets + hets2) %>%
      select(bin_id, beagle_s, eagle_s, shapeit_s, beagle_f, eagle_f, shapeit_f, hets)
  }
  return(final_df)
}

df_eur <- get_errors_mb(1:200, 1e6, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

df_eur %>%
  ggplot(aes(x = bin_id, y = hets)) +
  geom_point() +
  geom_line() +
  ggtitle("Aggregated Heterozygous Sites per Bin", "EUR Synthetic Diploids") +
  xlab("Bin") +
  ylab("Heterozygous Sites") +
  theme_classic()

df_eur %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "EUR Synthetic Diploids")

df_eur %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
  geom_smooth() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "EUR Synthetic Diploids")

df_eur %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_eur %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "spearman") %>%
  knitr::kable()

```

#### Adjust for het sites

```{r}
df_eur %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per Heterozygous Site per MB Bin", "EUR Synthetic Diploids")

df_eur %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_eur %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```


### Flips

```{r}
df_eur %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "EUR Synthetic Diploids")

df_eur %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
  geom_smooth() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "EUR Synthetic Diploids")

df_eur %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_eur %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_eur %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per Heterozygous Site per MB Bin", "EUR Synthetic Diploids")

df_eur %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_eur %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

### Higher Resolution

```{r}
df_eur <- get_errors_mb(1:200, 1e5, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

df_eur %>%
  ggplot(aes(x = bin_id, y = hets)) +
  geom_point() +
  geom_line() +
  ggtitle("Aggregated Heterozygous Sites per Bin", "EUR Synthetic Diploids") +
  xlab("Bin") +
  ylab("Heterozygous Sites") +
  theme_classic()

df_eur %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per 100KB Bin", "EUR Synthetic Diploids")

df_eur %>%
  mutate(eagle_beagle = eagle_s - beagle_s,
         eagle_shapeit = eagle_s - shapeit_s) %>%
  select(bin_id, eagle_beagle, eagle_shapeit) %>%
  pivot_longer(-bin_id, names_to = "Comparison", values_to = "Difference") %>%
  ggplot(aes(x = bin_id, y = Difference, color = Comparison, group = Comparison)) +
    geom_point() +
    theme_classic() +
    xlab("Bin") +
  geom_hline(yintercept = 0) +
  ggtitle("Difference in Switches per 100KB Bin", "EUR Synthetic Diploids") +
  geom_smooth(linewidth = 2.5, color = "black", se = FALSE)+
  geom_smooth()
  
df_eur %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per 100KB Bin", "EUR Synthetic Diploids")
```


## AFR

### Switches 

```{r}
df_afr <- get_errors_mb(201:400, 1e6, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

df_afr %>%
  ggplot(aes(x = bin_id, y = hets)) +
  geom_point() +
  geom_line() +
  ggtitle("Aggregated Heterozygous Sites per Bin", "AFR Synthetic Diploids") +
  xlab("Bin") +
  ylab("Heterozygous Sites") +
  theme_classic()

df_afr %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "AFR Synthetic Diploids")

df_afr %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  geom_smooth() +
  ggtitle("Switches per MB Bin", "AFR Synthetic Diploids")

df_afr %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_afr %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_afr %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per Heterozygous Site per MB Bin", "AFR Synthetic Diploids")

df_afr %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_afr %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

### Flips

```{r}
df_afr %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "AFR Synthetic Diploids")

df_afr %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "AFR Synthetic Diploids") +
  geom_smooth()

df_afr %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_afr %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_afr %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per Heterozygous Site per MB Bin", "AFR Synthetic Diploids")

df_afr %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_afr %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

## AMR

### Switches

```{r}
df_amr <- get_errors_mb(401:500, 1e6, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

df_amr %>%
  ggplot(aes(x = bin_id, y = hets)) +
  geom_point() +
  geom_line() +
  ggtitle("Aggregated Heterozygous Sites per Bin", "AMR Synthetic Diploids") +
  xlab("Bin") +
  ylab("Heterozygous Sites") +
  theme_classic()

df_amr %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "AMR Synthetic Diploids")

df_amr %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "AMR Synthetic Diploids") +
  geom_smooth()

df_amr %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_amr %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_amr %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per Heterozygous Site per MB Bin", "AMR Synthetic Diploids")

df_amr %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_amr %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

### Flips

```{r}
df_amr %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "AMR Synthetic Diploids")

df_amr %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "AMR Synthetic Diploids") +
  geom_smooth()

df_amr %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_amr %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_amr %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per Heterozygous Site per MB Bin", "AMR Synthetic Diploids")

df_amr %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_amr %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

## EAS

### Switches

```{r}
df_eas <- get_errors_mb(501:600, 1e6, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

df_eas %>%
  ggplot(aes(x = bin_id, y = hets)) +
  geom_point() +
  geom_line() +
  ggtitle("Aggregated Heterozygous Sites per Bin", "EAS Synthetic Diploids") +
  xlab("Bin") +
  ylab("Heterozygous Sites") +
  theme_classic()

df_eas %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "EAS Synthetic Diploids")

df_eas %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "EAS Synthetic Diploids") +
  geom_smooth()

df_eas %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_eas %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_eas %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per Heterozygous Site per MB Bin", "EAS Synthetic Diploids")

df_eas %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_eas %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

### Flips

```{r}
df_eas %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "EAS Synthetic Diploids")

df_eas %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "EAS Synthetic Diploids") +
  geom_smooth()

df_eas %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_eas %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_eas %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per Heterozygous Site per MB Bin", "EAS Synthetic Diploids")

df_eas %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_eas %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

## SAS

### Switches

```{r}
df_sas <- get_errors_mb(601:700, 1e6, het_loc_dir, beagle_switch_dir, eagle_switch_dir, shapeit_switch_dir, df_freq)

df_sas %>%
  ggplot(aes(x = bin_id, y = hets)) +
  geom_point() +
  geom_line() +
  ggtitle("Aggregated Heterozygous Sites per Bin", "SAS Synthetic Diploids") +
  xlab("Bin") +
  ylab("Heterozygous Sites") +
  theme_classic()

df_sas %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "SAS Synthetic Diploids")

df_sas %>%
  select(bin_id, beagle_s, eagle_s, shapeit_s) %>%
  rename(Beagle = beagle_s, Eagle = eagle_s, SHAPEIT = shapeit_s) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per MB Bin", "SAS Synthetic Diploids") +
  geom_smooth()

df_sas %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_sas %>%
  select(beagle_s, eagle_s, shapeit_s) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_sas %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Switches") %>%
  ggplot(aes(x = bin_id, y = Switches, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Switches per Heterozygous Site per MB Bin", "SAS Synthetic Diploids")

df_sas %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_sas %>%
  mutate(beagle = beagle_s / hets,
         eagle = eagle_s / hets,
         shapeit = shapeit_s / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

### Flips

```{r}
df_sas %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "SAS Synthetic Diploids")

df_sas %>%
  select(bin_id, beagle_f, eagle_f, shapeit_f) %>%
  rename(Beagle = beagle_f, Eagle = eagle_f, SHAPEIT = shapeit_f) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per MB Bin", "SAS Synthetic Diploids") +
  geom_smooth()

df_sas %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "pearson") %>%
  knitr::kable()

df_sas %>%
  select(beagle_f, eagle_f, shapeit_f) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

#### Adjust for het sites

```{r}
df_sas %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(bin_id, beagle, eagle, shapeit) %>%
  rename(Beagle = beagle, Eagle = eagle, SHAPEIT = shapeit) %>%
  pivot_longer(-bin_id, names_to = "Method", values_to = "Flips") %>%
  ggplot(aes(x = bin_id, y = Flips, color = Method )) +
    geom_point() +
    geom_line() +
    theme_classic() +
    xlab("Bin") +
  ggtitle("Flips per Heterozygous Site per MB Bin", "SAS Synthetic Diploids")

df_sas %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor() %>%
  knitr::kable()

df_sas %>%
  mutate(beagle = beagle_f / hets,
         eagle = eagle_f / hets,
         shapeit = shapeit_f / hets) %>%
  select(beagle, eagle, shapeit) %>%
  cor(method = "spearman") %>%
  knitr::kable()
```

